---
title: "Mako hSDM BRT preliminary results"
author: "Emily Nazario"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
code-block-bg: true
---

# Study hypotheses and objectives

**H1:** The AGI model will perform better than the dissolved oxygen and null model, and the dissolved oxygen model will perform better than the null model.

*study objective being met:* Which model performs the best and presents the best predictions (i.e., best predictive performance scores, most ecologically realistic suitability maps)?

**H2:** The inclusion of dissolved oxygen at deeper depths will result in better/more ecologically realistic habitat suitability predictions relative to the dissolved oxygen model considering surface values alone.

*study objective being met:* How does dissolved oxygen at different depths influence habitat suitability predictions relative to oxygen at the surface?

**H3:** The inclusion of the AGI at deeper depths will result in better/more ecologically realistic habitat suitability predictions relative to the AGI model considering surface values alone.

*study objective being met:* How does the aerobic growth index (AGI; environmental oxygen supply:theoretical oxygen demand) at different depths influence habitat suitability predictions relative to the aerobic growth index at the surface?

**H4:** There will be important relationships between dissolved oxygen/the AGI and latitude/distance to coast.

*study objective being met:* Are there any important relationships between dissolved oxygen or AGI at the surface or at depth and latitude or distance to the coast?

**H5:** The null model will predict higher habitat suitability in areas or during seasons or periods (upwelling or La Niña) with lower dissolved oxygen through the water column relative to the dissolved oxygen and AGI models.

*study objective being met:* How do the habitat suitability maps differ between the models? How do these variations compare for different points in time?

```{r}
#| label: libraries and data
#| echo: false
#| warning: false
#| message: false
#| include: false

#libraries
library(tidyverse)
library(gbm)
library(dismo)
library(here)
library(terra)
library(sf)
library(tidyterra)
library(here);here <- here::here #plyr's here function masks here::here
library(tidyquant)

set.seed(1004)

source(here("functions/BRT_evaluation_functions.R"))
source(here("functions/brt_explore_quarto_function.R"))
source(here("functions/hsi_rast_functions.R"))

# CRW outputs and data -------------------------------------------------------
brt_outputs_crw <- list.files(here("data/brt/mod_outputs/crw"), full.names = TRUE, pattern = ".rds")

#load test datasets
base_test_daily_crw <- readRDS(here("data/brt/mod_eval/base_test_daily.rds"))
base_test_seasonal_crw <- readRDS(here("data/brt/mod_eval/base_test_seasonal.rds"))
base_test_annual_crw <- readRDS(here("data/brt/mod_eval/base_test_annual.rds"))

do_test_daily_crw <- readRDS(here("data/brt/mod_eval/do_test_daily.rds"))
agi_test_daily_crw <- readRDS(here("data/brt/mod_eval/agi_test_daily.rds"))

do_test_daily_seasonal_annual_crw <- readRDS(here("data/brt/mod_eval/do_test_daily_seasonal_annual.rds"))
agi_test_daily_seasonal_annual_crw <- readRDS(here("data/brt/mod_eval/agi_test_daily_seasonal_annual.rds"))

# Background outputs and data ----------------------------------------------
brt_outputs_back <- list.files(here("data/brt/mod_outputs/background"), full.names = TRUE, pattern = ".rds")

#load test datasets
base_test_daily_back <- readRDS(here("data/brt/mod_eval/back/base_test_daily.rds"))
base_test_seasonal_back <- readRDS(here("data/brt/mod_eval/back/base_test_seasonal.rds"))
base_test_annual_back <- readRDS(here("data/brt/mod_eval/back/base_test_annual.rds"))

do_test_daily_back <- readRDS(here("data/brt/mod_eval/back/do_test_daily.rds"))
agi_test_daily_back <- readRDS(here("data/brt/mod_eval/back/agi_test_daily.rds"))

do_test_daily_seasonal_annual_back <- readRDS(here("data/brt/mod_eval/back/do_test_daily_seasonal_annual.rds"))
agi_test_daily_seasonal_annual_back <- readRDS(here("data/brt/mod_eval/back/agi_test_daily_seasonal_annual.rds"))

```

# Model summaries

::: panel-tabset
## Base model - crw

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_crw[7], 
            test_data = base_test_daily_crw)
```

## Base model - back

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_back[7], 
            test_data = base_test_daily_back)
```

## AGI overfit - crw

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_agi_0m_60m_250m_dail_seas_ann_no_wind.rds",
            test_data = agi_test_daily_seasonal_annual_crw)

```

## DO overfit - crw

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_do_0m_60m_250m_dail_seas_ann_no_wind.rds",
            test_data = do_test_daily_seasonal_annual_crw)

```

## AGI overfit - back

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/refined/brt_agi_0m_60m_250m_dail_seas_ann_no_wind.rds",
            test_data = agi_test_daily_seasonal_annual_back)

```

## DO overfit - back

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/refined/brt_do_0m_60m_250m_dail_seas_ann_no_wind.rds",
            test_data = do_test_daily_seasonal_annual_back)

```

## AGI final - crw

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_agi_0m_250m_dail_seas_ann.rds",
            test_data = agi_test_daily_seasonal_annual_crw)

```

## DO final - crw

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_do_0m_250m_dail_seas_ann.rds",
            test_data = do_test_daily_seasonal_annual_crw)

```

## AGI final - back

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/refined/brt_agi_0m_250m_dail_seas_ann.rds",
            test_data = agi_test_daily_seasonal_annual_back)

```

## DO final - back

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/refined/brt_do_0m_250m_dail_seas_ann.rds",
            test_data = do_test_daily_seasonal_annual_back)

```
:::

# Summary table of model performance metrics

```{r}
#| echo: false

output_sum <- read.csv(here("data/brt/mod_outputs/brt_prelim_results.csv"))
kableExtra::kable(output_sum)
```

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10
#| echo: false

ggplot(output_sum, aes(AUC, TSS, color = deviance_exp, label = model)) +
  geom_point(size = 5) +
  xlab('AUC') +
  ylab('TSS') +
  scale_color_gradientn(colors = MetBrewer::met.brewer("Greek")) +
  ggrepel::geom_label_repel(aes(label = model),
                  box.padding   = 0.35,
                  point.padding = 1,
                  segment.color = 'grey50',
                  max.overlaps = 40,
                  label.size = 0.5)+
  theme_minimal()+
  theme(legend.position = "right")

```

## Conclusions

-   Based on the above comparisons between models using PAs from CRW and background sampling approaches, I have decided to move forward using the CRW PA models. The background sampling models consistently indicated that they were overfit, and thus the CRW models are the preferred choice.

-   Given the similar values among performance metrics between the overfit models (those with all depth levels and temporal resolutions) and the models without data at 60m, we have selected the latter to be the final study models. We explored similar models without data at 60m and at a seasonal temporal resolution, but these models performed considerably worse. Additionally, DO and AGI at 0m and 250m at a seasonal resolution were often a predictor variable with a top 5 highest relative importance, thus, keeping data at a seasonal resolution seemed valuable.

# Final models w/o random predictor variables

::: panel-tabset
## Base model

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/final_mods/brt_base_0m_dail_no_wind.rds",
            test_data = base_test_daily_crw)

```

## DO overfit model

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/final_mods/brt_do_0m_60m_250m_dail_seas_ann_no_wind.rds",
            test_data = do_test_daily_seasonal_annual_crw)

```

## AGI overfit model

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/final_mods/brt_agi_0m_60m_250m_dail_seas_ann_no_wind.rds",
            test_data = agi_test_daily_seasonal_annual_crw)

```

## DO final model

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/final_mods/brt_do_0m_250m_dail_seas_ann.rds",
            test_data = do_test_daily_seasonal_annual_crw)

```

## AGI final model

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/final_mods/brt_agi_0m_250m_dail_seas_ann.rds",
            test_data = agi_test_daily_seasonal_annual_crw)

```
:::

# HSI maps

## List models

```{r}
#| echo: false
#| warning: false
#| include: false

base_mod <- readRDS(here("data/brt/mod_outputs/final_mods/brt_base_0m_dail_no_wind.rds"))
do_mod_of <- readRDS(here("data/brt/mod_outputs/final_mods/brt_do_0m_60m_250m_dail_seas_ann_no_wind.rds"))
do_mod_fin <- readRDS(here("data/brt/mod_outputs/final_mods/brt_do_0m_250m_dail_seas_ann.rds"))
agi_mod_of <- readRDS(here("data/brt/mod_outputs/final_mods/brt_agi_0m_60m_250m_dail_seas_ann_no_wind.rds"))
agi_mod_fin <- readRDS(here("data/brt/mod_outputs/final_mods/brt_agi_0m_250m_dail_seas_ann.rds"))

```

::: panel-tabset
## Average across study period

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

#hsi_rast_gen(date_start = c("2003-01-01"), date_end = c("2015-12-31"), season = "WSpSuF", output_name = "Jan03_Dec15")

hsi_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Jan03_Dec15")

```

## Average by ENSO phase

These dates run from January to December of each year, but I am happy to modify this if there are better ways to categorize/plot ENSO phases.

### Neutral year (2013)

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

#hsi_rast_gen(date_start = c("2013-01-01"), date_end = c("2013-12-31"), season = "WSpSuF", output_name = "Jan13_Dec13")

hsi_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Jan13_Dec13")

```

### Strong La Niña year (-1.5 to -1.9 temperature anomaly; 2011)

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

#hsi_rast_gen(date_start = c("2011-01-01"), date_end = c("2011-12-31"), season = "WSpSuF", output_name = "Jan11_Dec11")

hsi_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Jan11_Dec11")

```

### Moderate/strong El Niño year (1.5 to 1.9 temperature anomaly; 2009)

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

#hsi_rast_gen(date_start = c("2009-01-01"), date_end = c("2009-12-31"), season = "WSpSuF", output_name = "Jan09_Dec09")

hsi_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Jan09_Dec09")

```

## Random day (September 24, 2007)

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

#hsi_rast_gen(date_start = c("2007-09-24"), date_end = c("2007-09-24"), season = "F", output_name = "Sept_24_2007")

hsi_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Sept_24_2007")

```
:::

# HSI difference maps

::: panel-tabset
## Average across study period

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

hsi_diff_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Jan03_Dec15")

```

## Average by ENSO phase

### Neutral year (2013)

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

hsi_diff_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Jan13_Dec13")

```

### Strong La Niña year (-1.5 to -1.9 temperature anomaly; 2011)

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

hsi_diff_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Jan11_Dec11")

```

### Moderate/strong El Niño year (1.5 to 1.9 temperature anomaly; 2009)

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

hsi_diff_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Jan09_Dec09")

```

## Random day (September 24, 2007)

```{r}
#| echo: false
#| warning: false
#| fig-width: 12
#| fig-height: 5

hsi_diff_maps(rast_folder = "data/enviro/psat_spot_all/hsi_rasts/Sept_24_2007")

```
:::

# AGI maps

# Predictor variable relative importance

# Performance metric comparisons
