---
title: "Mako hSDM BRT explore (Background PAs)"
author: "Emily Nazario"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
code-block-bg: true
---

On this document, I've included the results from the initial exploration into the different model outputs, ranking of covariate influence, performance metrics, and prediction maps. This first set of models only includes extracted covariate data at a daily temporal resolution, but I am also considering exploring models that include covariate data at a seasonal or annual temporal resolution. The pseudo absences used in these models were generated using background sampling approaches. Lastly, hyperparameters were tuned using the caret package and across all models, a learning rate of 0.05 and tree complexity of 3 resulted in the highest accuracy. Lastly, the 'pred_var' predictor is a random set of numbers that will be used to identify which predictor variables should be included in the final model, and which are not informative.

The hypotheses I would like to test with these models are as follows:

**H1:** The AGI model will perform better than the dissolved oxygen and null model, and the dissolved oxygen model will perform better than the null model.

*study objective being met:* Which model performs the best and presents the best predictions (i.e., best predictive performance scores, most ecologically realistic suitability maps)?

**H2:** The inclusion of dissolved oxygen at deeper depths will result in better/more ecologically realistic habitat suitability predictions relative to the dissolved oxygen model considering surface values alone.

*study objective being met:* How does dissolved oxygen at different depths influence habitat suitability predictions relative to oxygen at the surface?

**H3:** The inclusion of the AGI at deeper depths will result in better/more ecologically realistic habitat suitability predictions relative to the AGI model considering surface values alone.

*study objective being met:* How does the aerobic growth index (AGI; environmental oxygen supply:theoretical oxygen demand) at different depths influence habitat suitability predictions relative to the aerobic growth index at the surface?

**H4:** There will be important relationships between dissolved oxygen/the AGI and latitude/distance to coast.

*study objective being met:* Are there any important relationships between dissolved oxygen or AGI at the surface or at depth and latitude or distance to the coast?

**H5:** The null model will predict higher habitat suitability in areas or during seasons or periods (upwelling or La Niña) with lower dissolved oxygen through the water column relative to the dissolved oxygen and AGI models.

*study objective being met:* How do the habitat suitability maps differ between the models? How do these variations compare for different points in time?

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

#libraries
library(tidyverse)
library(gbm)
library(dismo)
library(here)
library(terra)
library(sf)
library(tidyterra)
library(here);here <- here::here #plyr's here function masks here::here

set.seed(1004)

source(here("functions/BRT_evaluation_functions.R"))
source(here("functions/brt_explore_quarto_function.R"))

#load BRT outputs
brt_outputs <- list.files(here("data/brt/mod_outputs/background"), full.names = TRUE, pattern = ".rds")
brt_outputs_Ntag <- list.files(here("data/brt/mod_outputs/background/no_tag"), full.names = TRUE)

#load test datasets
base_test_daily <- readRDS(here("data/brt/mod_eval/back/base_test_daily.rds"))
base_test_seasonal <- readRDS(here("data/brt/mod_eval/back/base_test_seasonal.rds"))
base_test_annual <- readRDS(here("data/brt/mod_eval/back/base_test_annual.rds"))

do_test_daily <- readRDS(here("data/brt/mod_eval/back/do_test_daily.rds"))
agi_test_daily <- readRDS(here("data/brt/mod_eval/back/agi_test_daily.rds"))

do_test_daily_seasonal_annual <- readRDS(here("data/brt/mod_eval/back/do_test_daily_seasonal_annual.rds"))
agi_test_daily_seasonal_annual <- readRDS(here("data/brt/mod_eval/back/agi_test_daily_seasonal_annual.rds"))

```

# Base models

These three models represent three different options for the base model and either include spatial predictors, a tag ID predictor, both, or neither. These models were developed by splitting the data set into 75/25 train/test, and thus that is the model evaluation approach used here. However, once a model is selected, I can run additional evaluation metrics (i.e., LOO, k-fold). I can also complete these now depending on when that is typically performed.

::: panel-tabset
## 0m, no spatial predictors, no tag ID predictor

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[7], 
            test_data = base_test_daily)

```

## 0m, no spatial predictors, yes tag ID predictor

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[8], 
            test_data = base_test_daily)

```

## 0m, yes spatial predictors, yes tag ID predictor

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[9], 
            test_data = base_test_daily)

```
:::

# DO models

I ran a suite of models that include various combinations of data at depth, spatial predictors, and tag ID predictors. Moving forward, I would also like to include DO and the other environmental predictor variables as longer time scales (seasonal/annual).

::: panel-tabset
## 0m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[14], 
            test_data = do_test_daily)

```

## 0m, yes spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[15], 
            test_data = do_test_daily)

```

## 0m & 60m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[13], 
            test_data = do_test_daily)

```

## 0m & 250m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[10], 
            test_data = do_test_daily)

```

## 0m, 60m, & 250m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[11], 
            test_data = do_test_daily)

```

## 0m, 60m, & 250m, yes spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[12], 
            test_data = do_test_daily)

```
:::

# AGI models

I ran a suite of models that include various combinations of data at depth, spatial predictors, and tag ID predictors. Moving forward, I would also like to include AGI and the other environmental predictor variables as longer time scales (seasonal/annual).

::: panel-tabset
## 0m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[5], 
            test_data = agi_test_daily)

```

## 0m, yes spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[6], 
            test_data = agi_test_daily)

```

## 0m & 60m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[4], 
            test_data = agi_test_daily)

```

## 0m & 250m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[1], 
            test_data = agi_test_daily)

```

## 0m, 60m, & 250m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[2], 
            test_data = agi_test_daily)

```

## 0m, 60m, & 250m, yes spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[3], 
            test_data = agi_test_daily)

```
:::

# Summary table of results

```{r}
#| warning: false
#| message: false

output_sum <- read.csv(here("data/brt/mod_outputs/brt_bckg_output_summary.csv"))
kableExtra::kable(output_sum)

```

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

ggplot(output_sum, aes(AUC, TSS, color = deviance_exp, label = model)) +
  geom_point(size = 5) +
  xlab('AUC') +
  ylab('TSS') +
  scale_color_gradientn(colors = MetBrewer::met.brewer("Greek")) +
  ggrepel::geom_label_repel(aes(label = model),
                  box.padding   = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50',
                  max.overlaps = 20,
                  label.size = 0.5)

```

## Conclusions from initial models w/ tag ID

-   **Base models:** Relative to the CRW PA base models, these had drastically higher AUC scores and deviance explained values. The base model with no spatial or tag ID predictors was the lowest scoring model.

-   DO and AGIModel performance generally increased with the added depth layers, but were all fairly comparable to each other. Models with spatial and tag ID predictors performed the best, but as described on the CRW PA document, we will likely not include them for these models as they would not be included in the projection work and are not essential for addressing this study's objectives.

-   The performance metrics across comparable DO and AGI models were much more similar relative to the models with the CRW PA data.

# DO models w/o tag ID

Here, I have run the same models as above, but without tag ID as a predictor variable. For this chunk of models, I am interested in identifying the role that dissolved oxygen may play in habitat suitability predictions, and how its relative importance compares to other covariates that are typically included in SDMs. Additionally, as BRTs are nonparametric, it is not critical or necessary for tag ID to be included.

::: panel-tabset
## 0m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[12], 
            test_data = do_test_daily)

```

## 0m, yes spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[13], 
            test_data = do_test_daily)

```

## 0m & 60m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[11], 
            test_data = do_test_daily)

```

## 0m & 250m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[8], 
            test_data = do_test_daily)

```

## 0m, 60m, & 250m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[9], 
            test_data = do_test_daily)

```

## 0m, 60m, & 250m, yes spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[10], 
            test_data = do_test_daily)

```
:::

# AGI models w/o tag ID

Here, I have run the same models as above, but without tag ID as a predictor variable. For this chunk of models, I am interested in identifying the role that AGI may play in habitat suitability predictions, and how its relative importance compares to other covariates that are typically included in SDMs. Additionally, as BRTs are nonparametric, it is not critical or necessary for tag ID to be included.

::: panel-tabset
## 0m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[5], 
            test_data = agi_test_daily)

```

## 0m, yes spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[6], 
            test_data = agi_test_daily)

```

## 0m & 60m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[4], 
            test_data = agi_test_daily)

```

## 0m & 250m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[1], 
            test_data = agi_test_daily)
```

## 0m, 60m, & 250m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[2], 
            test_data = agi_test_daily)

```

## 0m, 60m, & 250m, yes spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[3], 
            test_data = agi_test_daily)

```
:::

# Summary table of results

```{r}
#| warning: false
#| message: false

output_sum_Ntag <- read.csv(here("data/brt/mod_outputs/brt_bckg_output_summary_Ntag.csv"))
kableExtra::kable(output_sum_Ntag)

```

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

output_sum_Ntag_Nspat <- output_sum_Ntag %>% 
  filter(!grepl("Yspat", model))

ggplot(output_sum_Ntag_Nspat, aes(AUC, TSS, color = deviance_exp, label = model)) +
  geom_point(size = 5) +
  xlab('AUC') +
  ylab('TSS') +
  scale_color_gradientn(colors = MetBrewer::met.brewer("Greek")) +
  ggrepel::geom_label_repel(aes(label = model),
                  box.padding   = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50',
                  max.overlaps = 20,
                  label.size = 0.5)

```

## Conclusions from initial models w/o tag ID

-   If only considering models that did not include spatial data as model predictors, the AGI models performed much better than the DO models across the board.

-   The AGI model will all depth layers performed the best and considerably better than the comparable DO model.

-   For the DO model with all depth layers, DO_0m was the predictor variable with the highest relative influence, but was closely followed by bathymetry. DO_250m was the third most influential predictor, but is considerably lower than DO_0m and bathymetry. Partial plots show drastically different relationships that the CRW PA models, with DO_250m having a positive correlation and DO_0m having an inverse sweet spot.

-   For the AGI model with all depth layers, bathymetry and temperature were the two predictors with the highest relative influence, and AGI 250m was listed third, somewhat closely followed by AGI 0m. The partial plots for these two variables are similar to the DO models, but less extreme.

# Base models w/o tag ID and w/ data at seasonal and annual resolutions

For these models, the environmental raster data was averaged according to season and year. Observed and pseudo absence locations were then used for environmental data extraction along these raster files and were matched to each file according to either the season or year.

::: panel-tabset
## Seasonal resolution

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/seasonal/brt_base_0m_seas_Nspat_Ntag.rds",
            test_data = base_test_seasonal)

```

## Annual resolution

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/annual/brt_base_0m_ann_Nspat_Ntag.rds",
            test_data = base_test_annual)
```
:::

# DO models w/o tag ID and w/ data at seasonal and annual resolutions

::: panel-tabset
## Seasonal, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/seasonal/brt_do_0m_60m_250m_seas_Nspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Seasonal, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/seasonal/brt_do_0m_60m_250m_seas_Yspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Annual, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/annual/brt_do_0m_60m_250m_ann_Nspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Annual, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/annual/brt_do_0m_60m_250m_ann_Yspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Daily, seasonal, and Annual, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/annual/brt_do_0m_60m_250m_dail_seas_ann_Nspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Daily, Seasonal, and Annual, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/annual/brt_do_0m_60m_250m_dail_seas_ann_Yspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```
:::

# AGI models w/o tag ID and w/ data at seasonal and annual resolutions

::: panel-tabset
## Seasonal, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/seasonal/brt_agi_0m_60m_250m_seas_Nspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Seasonal, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/seasonal/brt_agi_0m_60m_250m_seas_Yspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Annual, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/annual/brt_agi_0m_60m_250m_ann_Nspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Annual, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/annual/brt_agi_0m_60m_250m_ann_Yspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Daily, seasonal, and Annual, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/annual/brt_agi_0m_60m_250m_dail_seas_ann_Nspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Daily, Seasonal, and Annual, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/background/annual/brt_agi_0m_60m_250m_dail_seas_ann_Yspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```
:::

# Summary table of results

```{r}
#| warning: false
#| message: false

output_sum_seas_ann <- read.csv(here("data/brt/mod_outputs/brt_background_seas_ann_output_summary.csv"))
kableExtra::kable(output_sum)
```

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

output_sum_seas_ann_Nspat <- output_sum_seas_ann %>% 
  filter(!grepl("Yspat", model))

ggplot(output_sum_seas_ann_Nspat, aes(AUC, TSS, color = deviance_exp, label = model)) +
  geom_point(size = 5) +
  xlab('AUC') +
  ylab('TSS') +
  scale_color_gradientn(colors = MetBrewer::met.brewer("Greek")) +
  ggrepel::geom_label_repel(aes(label = model),
                  box.padding   = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50',
                  max.overlaps = 20,
                  label.size = 0.5)

```

## Conclusions from initial seasonal/annual models 

-   Seasonal and annual base models performed better than the daily resolution base models, with the annual base model performing better than the seasonal one.

-   The DO and AGI models with all depth layers and temporal resolutions were by far the best performing and had nearly identical scores across evaluation metrics. The models that also included spatial predictors also performed slightly better than those without, but were still fairly comparable.

-   For the DO model with all temporal resolutions, the top predictor variables with the highest relative importance were bathymetry and DO_0m_daily. The next variables that have considerably lower values are DO_250m_seasonal and DO_0m_seasonal. Partial plots follow similar trends as previously described.

-   For the AGI model with all temporal resolutions, bathymetry and temperature were the two predictors with the highest relative influence. The next variables that have considerably lower values are AGI_250m_seasonal and AGI_0m_seasonal.
