---
title: "Mako hSDM BRT explore (CRW PAs)"
author: "Emily Nazario"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
code-block-bg: true
---

On this document, I've included the results from the initial exploration into the different model outputs, ranking of covariate influence, performance metrics, and prediction maps.

The majority of the predictors included in the following models are at a daily temporal resolution. However, for the DO and AGI models, we also investigated the inclusion of these two predictors at seasonal and annual temporal resolutions. The remaining environmental predictors are also available at these resolutions, and can included in follow-up models.

The pseudo absences used in these models were generated using correlated random walk approaches, but another quarto document includes models with background sampling pseudo absences. Lastly, hyperparameters were tuned using the caret package and across all models, a learning rate of 0.05 and tree complexity of 3 resulted in the highest accuracy. Lastly, the 'pred_var' predictor is a random set of numbers that will be used to identify which predictor variables should be included in the final model, and which are not informative.

The hypotheses I would like to test with these models are as follows:

**H1:** The AGI model will perform better than the dissolved oxygen and null model, and the dissolved oxygen model will perform better than the null model.

*study objective being met:* Which model performs the best and presents the best predictions (i.e., best predictive performance scores, most ecologically realistic suitability maps)?

**H2:** The inclusion of dissolved oxygen at deeper depths will result in better/more ecologically realistic habitat suitability predictions relative to the dissolved oxygen model considering surface values alone.

*study objective being met:* How does dissolved oxygen at different depths influence habitat suitability predictions relative to oxygen at the surface?

**H3:** The inclusion of the AGI at deeper depths will result in better/more ecologically realistic habitat suitability predictions relative to the AGI model considering surface values alone.

*study objective being met:* How does the aerobic growth index (AGI; environmental oxygen supply:theoretical oxygen demand) at different depths influence habitat suitability predictions relative to the aerobic growth index at the surface?

**H4:** There will be important relationships between dissolved oxygen/the AGI and latitude/distance to coast.

*study objective being met:* Are there any important relationships between dissolved oxygen or AGI at the surface or at depth and latitude or distance to the coast?

**H5:** The null model will predict higher habitat suitability in areas or during seasons or periods (upwelling or La Niña) with lower dissolved oxygen through the water column relative to the dissolved oxygen and AGI models.

*study objective being met:* How do the habitat suitability maps differ between the models? How do these variations compare for different points in time?

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

#libraries
library(tidyverse)
library(gbm)
library(dismo)
library(here)
library(terra)
library(sf)
library(tidyterra)
library(here);here <- here::here #plyr's here function masks here::here
library(plotly)

set.seed(1004)

source(here("functions/BRT_evaluation_functions.R"))
source(here("functions/brt_explore_quarto_function.R"))

#load BRT outputs (daily tag and no tag files only)
brt_outputs <- list.files(here("data/brt/mod_outputs/crw"), full.names = TRUE, pattern = ".rds")
brt_outputs_Ntag <- list.files(here("data/brt/mod_outputs/crw/no_tag"), full.names = TRUE)

#load test datasets
base_test_daily <- readRDS(here("data/brt/mod_eval/base_test_daily.rds"))
base_test_seasonal <- readRDS(here("data/brt/mod_eval/base_test_seasonal.rds"))
base_test_annual <- readRDS(here("data/brt/mod_eval/base_test_annual.rds"))

do_test_daily <- readRDS(here("data/brt/mod_eval/do_test_daily.rds"))
agi_test_daily <- readRDS(here("data/brt/mod_eval/agi_test_daily.rds"))

do_test_daily_seasonal_annual <- readRDS(here("data/brt/mod_eval/do_test_daily_seasonal_annual.rds"))
agi_test_daily_seasonal_annual <- readRDS(here("data/brt/mod_eval/agi_test_daily_seasonal_annual.rds"))

```

# Base models

These three models represent three different options for the base model and either include spatial predictors, a tag ID predictor, both, or neither. These models were developed by splitting the data set into 75/25 train/test, and thus that is the model evaluation approach used here. However, once a model is selected, I can run additional evaluation metrics (i.e., LOO, k-fold). I can also complete these now depending on when that is typically performed.

::: panel-tabset
## 0m, no spatial predictors, no tag ID predictor

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[7], 
            test_data = base_test_daily)

```

## 0m, no spatial predictors, yes tag ID predictor

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[8], 
            test_data = base_test_daily)
```

## 0m, yes spatial predictors, yes tag ID predictor

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[9], 
            test_data = base_test_daily)

```
:::

# DO models

I ran a suite of models that include various combinations of data at depth, spatial predictors, and tag ID predictors. Moving forward, I would also like to include DO and the other environmental predictor variables as longer time scales (seasonal/annual).

::: panel-tabset
## 0m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[14], 
            test_data = do_test_daily)

```

## 0m, yes spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[15], 
            test_data = do_test_daily)

```

## 0m & 60m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[13], 
            test_data = do_test_daily)

```

## 0m & 250m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[10], 
            test_data = do_test_daily)

```

## 0m, 60m, & 250m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[11], 
            test_data = do_test_daily)

```

## 0m, 60m, & 250m, yes spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[12], 
            test_data = do_test_daily)

```
:::

# AGI models

I ran a suite of models that include various combinations of data at depth, spatial predictors, and tag ID predictors. Moving forward, I would also like to include AGI and the other environmental predictor variables as longer time scales (seasonal/annual).

::: panel-tabset
## 0m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[5], 
            test_data = agi_test_daily)

```

## 0m, yes spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[6], 
            test_data = agi_test_daily)

```

## 0m & 60m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[4], 
            test_data = agi_test_daily)

```

## 0m & 250m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[1], 
            test_data = agi_test_daily)

```

## 0m, 60m, & 250m, no spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[2], 
            test_data = agi_test_daily)

```

## 0m, 60m, & 250m, yes spatial, yes tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs[3], 
            test_data = agi_test_daily)

```
:::

# Summary table of results

```{r}
#| warning: false
#| message: false

output_sum <- read.csv(here("data/brt/mod_outputs/brt_crw_output_summary.csv"))
kableExtra::kable(output_sum)

```

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

ggplot(output_sum, aes(x = AUC, y = TSS, color = deviance_exp, text = model)) +
  geom_point(size = 5) +
  xlab('AUC') +
  ylab('TSS') +
  scale_color_gradientn(colors = MetBrewer::met.brewer("Greek")) +
  ggrepel::geom_label_repel(aes(label = model),
                  box.padding   = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50', 
                  max.overlaps = 20,
                  label.size = 0.5)
```

## Conclusions from initial models w/ tag ID

-   **Base models:**Bathymetry was consistently one of the top predictor variables across all base models, and percent explained greatly increased after including spatial and tag ID as additional predictors. After running these initial models, we decided to instead run the spatial analysis separately (GLMs, GAMs), rather than including them as predictors in the hSDMs, to specifically investigate the relationships between latitude, distance to coast, and the AGI or DO at different depth layers. Additionally, we will not include tag ID as a predictor variable as it would not be included in any projection work and is not critical for the main objectives of this study.

-   **DO models:** Performance metrics generally increased, though only subtly, after including the additional depth layers relative to the DO_0m model. However, relative to the base models, including DO considerably improved model performance. Across depth layers, DO at 0m and 250m were consistently in the top 5 predictors with most relative influence and had comparable contributions. From the partial plots, we generally see a sweet spot for DO values at 0m and a negative relationship for DO at 250m.

-   **AGI models:** Performance metrics were comparable among the DO and AGI models, and the patterns observed for the DO models also generally held for the AGI models. We see model performance greatly improve for the AGI models relative to the base models, and performance also subtly increased after including the additional depth layers. A primary difference for the AGI models is the relative influence of the AGI at 250m. For these models, the AGI at this depth layer is the only one appearing in the top variables with the highest relative influence, and the AGI at 0m and 60m is typically lower in the list. The AGI partial plots show similar patterns as the DO plots, with less of a dramatic negative relationship for the AGI at 250m.

-   The random predictor variable was typically the lowest performing metric, but across some models, had a higher relative influence than the predictors related to wind stress and wind stress curl.

# DO models w/o tag ID

Here, I have run the same models as above, but without tag ID as a predictor variable. For this chunk of models, I am interested in identifying the role that dissolved oxygen may play in habitat suitability predictions, and how its relative importance compares to other covariates that are typically included in SDMs. Additionally, as BRTs are nonparametric, it is not critical or necessary for tag ID to be included.

::: panel-tabset
## 0m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[12], 
            test_data = do_test_daily)

```

## 0m, yes spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[13], 
            test_data = do_test_daily)

```

## 0m & 60m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[11], 
            test_data = do_test_daily)

```

## 0m & 250m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[8], 
            test_data = do_test_daily)

```

## 0m, 60m, & 250m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[9], 
            test_data = do_test_daily)

```

## 0m, 60m, & 250m, yes spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[10], 
            test_data = do_test_daily)

```
:::

# AGI models w/o tag ID

Here, I have run the same models as above, but without tag ID as a predictor variable. For this chunk of models, I am interested in identifying the role that AGI may play in habitat suitability predictions, and how its relative importance compares to other covariates that are typically included in SDMs. Additionally, as BRTs are nonparametric, it is not critical or necessary for tag ID to be included.

::: panel-tabset
## 0m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[5], 
            test_data = agi_test_daily)

```

## 0m, yes spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[6], 
            test_data = agi_test_daily)

```

## 0m & 60m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[4], 
            test_data = agi_test_daily)

```

## 0m & 250m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[1], 
            test_data = agi_test_daily)

```

## 0m, 60m, & 250m, no spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[2], 
            test_data = agi_test_daily)

```

## 0m, 60m, & 250m, yes spatial, no tag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = brt_outputs_Ntag[3], 
            test_data = agi_test_daily)

```
:::

# Summary table of results

```{r}
#| warning: false
#| message: false

output_sum_Ntag <- read.csv(here("data/brt/mod_outputs/brt_crw_output_summary_Ntag.csv"))
kableExtra::kable(output_sum_Ntag)

```

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

output_sum_Ntag_Nspat <- output_sum_Ntag %>%
  filter(!grepl("Yspat", model))

ggplot(output_sum_Ntag_Nspat, aes(AUC, TSS, color = deviance_exp, label = model)) +
  geom_point(size = 5) +
  xlab('AUC') +
  ylab('TSS') +
  scale_color_gradientn(colors = MetBrewer::met.brewer("Greek")) +
  ggrepel::geom_label_repel(aes(label = model),
                  box.padding   = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50', 
                  max.overlaps = 20, 
                  label.size = 0.5)

```

## Conclusions from initial models w/o tag ID

-   These models were all developed using predictor data at a daily resolution, and if we only consider models with no spatial predictors, the DO model with DO data at 0m, 60m, and 250m performed the best, with the comparable AGI model having lower TSS and AUC scores.

-   The DO and AGI models both performed better relative to the base model.

-   DO at 0m and DO at 250m were the two predictors with the highest relative influence, while DO at 60m was considerably lower in the list. This pattern held whether or not spatial predictor variables were included. Still, performance metrics improved for the DO_0m_60m_250m model relative to the DO_0m_250m model. Partial plot patters for DO at 0m and 250m were the same as the original models that included tag ID as a predictor (sweet spot for 0m, negative relationship for 250m).

-   AGI at 250m was the most important predictor variable, followed by temperature and AGI at 0m (the two had nearly identical relative influence values). However, AGI at 0m became more influential if spatial predictors were included. Model performance had smaller differences between the AGI_0m_250m and AGI_0m_60m_250m as the DO models did. Partial plot patters for the AGI at 0m and 250m remained the same as described above.

# Base models w/o tag ID and w/ data at seasonal and annual resolutions

For these models, the environmental raster data was averaged according to season and year. Observed and pseudo absence locations were then used for environmental data extraction along these raster files and were matched to each file according to either the season or year.

::: panel-tabset
## Seasonal resolution

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/seasonal/brt_base_0m_seas_Nspat_Ntag.rds",
            test_data = base_test_seasonal)

```

## Annual resolution

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/annual/brt_base_0m_ann_Nspat_Ntag.rds",
            test_data = base_test_annual)
```
:::

# DO models w/o tag ID and w/ data at seasonal and annual resolutions

::: panel-tabset
## Seasonal, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/seasonal/brt_do_0m_60m_250m_seas_Nspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Seasonal, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/seasonal/brt_do_0m_60m_250m_seas_Yspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Annual, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/annual/brt_do_0m_60m_250m_ann_Nspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Annual, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/annual/brt_do_0m_60m_250m_ann_Yspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Daily, seasonal, and Annual, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/annual/brt_do_0m_60m_250m_dail_seas_ann_Nspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```

## Daily, Seasonal, and Annual, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/annual/brt_do_0m_60m_250m_dail_seas_ann_Yspat_Ntag.rds",
            test_data = do_test_daily_seasonal_annual)

```
:::

# AGI models w/o tag ID and w/ data at seasonal and annual resolutions

::: panel-tabset
## Seasonal, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/seasonal/brt_agi_0m_60m_250m_seas_Nspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Seasonal, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/seasonal/brt_agi_0m_60m_250m_seas_Yspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Annual, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/annual/brt_agi_0m_60m_250m_ann_Nspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Annual, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/annual/brt_agi_0m_60m_250m_ann_Yspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Daily, seasonal, and Annual, Nspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/annual/brt_agi_0m_60m_250m_dail_seas_ann_Nspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## Daily, Seasonal, and Annual, Yspat, Ntag

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/annual/brt_agi_0m_60m_250m_dail_seas_ann_Yspat_Ntag.rds",
            test_data = agi_test_daily_seasonal_annual)

```
:::

# Summary table of results

```{r}
#| warning: false
#| message: false

output_sum_seas_ann <- read.csv(here("data/brt/mod_outputs/brt_crw_seas_ann_output_summary.csv"))
kableExtra::kable(output_sum_seas_ann)
```

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

output_sum_seas_ann_Nspat <- output_sum_seas_ann %>%
  filter(!grepl("Yspat", model))

ggplot(output_sum_seas_ann_Nspat, aes(AUC, TSS, color = deviance_exp, label = model)) +
  geom_point(size = 5) +
  xlab('AUC') +
  ylab('TSS') +
  scale_color_gradientn(colors = MetBrewer::met.brewer("Greek")) +
  ggrepel::geom_label_repel(aes(label = model),
                  box.padding   = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50', 
                  max.overlaps = 20, 
                  label.size = 0.5)

```

## Conclusions from initial seasonal/annual models

-   Seasonal and annual base models were comparable in performance to the daily resolution base model, with seasonal performing slightly worse, and annual performing slightly better.

-   The AGI model with all depth layers and resolutions performed the best if only looking at models with no spatial predictor variables, but the comparable DO model performed similarly.

-   Annual models generally performed better than seasonal ones, but the models with data at a daily, seasonal, and annual data performed considerably better.

-   For the DO model with all depths and temporal resolutions, the two predictors with the highest relative influence (and whose values were quite comparable) were DO_250m_annual and DO_0m_daily. The remaining seasonal DO values were also highly ranked. Partial plots either show a negative correlation or a sweet spot range of DO values at each of the depth layers and resolutions.

-   For the AGI model with all depths and temporal resolutions, the top predictor variable is AGI_250m_annual, which is closely followed by daily temperature at 0m. Lower down the list is AGI_0m_daily, bathymetry, and AGI_0m_seasonal. Partial plot relationships show similar trends as described previously.

# Model fine-tuning and selection

Here, I take the two best performing models from the above sections (agi and do with all depths and temporal resolutions without tag ID or spatial variables as predictors) to be used as overfit reference models. The following model options excluded the wind predictors as these consistently had lower relative importance than the random predictor variable we included. I also included a combo model that uses information about AGI at 250 m and DO at 0m across temporal resolutions. Lastly, the final models also remove do/agi at 60m and at a seasonal resolution, as these were typically the vars with the lowest predictive performance relative to the other depth layers and resolutions.

::: panel-tabset

## Base no wind
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_base_0m_dail_no_wind.rds",
            test_data = base_test_daily)

```

## DO no wind
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_do_0m_60m_250m_dail_seas_ann_no_wind.rds",
            test_data = do_test_daily_seasonal_annual)

```

## AGI no wind
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_agi_0m_60m_250m_dail_seas_ann_no_wind.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## DO 0m, AGI 250m
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_agi_250_DO_0_dail_seas_ann.rds",
            test_data = readRDS(here("data/brt/mod_eval/agi_do_test_daily_seasonal_annual.rds")))

```

## DO w/o 60m 
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_do_0m_250m_dail_seas_ann.rds",
            test_data = do_test_daily_seasonal_annual)

```

## DO w/o seasonal
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_do_0m_60m_250m_dail_ann.rds",
            test_data = do_test_daily_seasonal_annual)

```

## DO w/o daily
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_do_0m_60m_250m_seas_ann.rds",
            test_data = do_test_daily_seasonal_annual)

```

## AGI w/o 60m 
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_agi_0m_250m_dail_seas_ann.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## AGI w/o seasonal
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_agi_0m_60m_250m_dail_ann.rds",
            test_data = agi_test_daily_seasonal_annual)

```

## AGI w/o daily
```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 10

explore_brt(mod_file_path = "data/brt/mod_outputs/crw/refined/brt_agi_0m_60m_250m_seas_ann.rds",
            test_data = agi_test_daily_seasonal_annual)

```

:::

# Summary table of results

```{r}
#| warning: false
#| message: false

output_sum_refined <- read.csv(here("data/brt/mod_outputs/brt_crw_refined_output_summary.csv"))
kableExtra::kable(output_sum_seas_ann)
```

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

ggplot(output_sum_refined, aes(AUC, TSS, color = deviance_exp, label = model)) +
  geom_point(size = 5) +
  xlab('AUC') +
  ylab('TSS') +
  scale_color_gradientn(colors = MetBrewer::met.brewer("Greek")) +
  ggrepel::geom_label_repel(aes(label = model),
                  box.padding   = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50', 
                  max.overlaps = 20, 
                  label.size = 0.5)

```

## Conclusions from adjusted models 

-   The DO or AGI annual values at 250 m and the DO or AGI daily values at 0m were consistently those with the highest relative importance. 
-   The reference models (that are likely overfit) still performed the best, with the AGI model having the highest scores across performance metrics. 
-   Seems like removing the wind predictors doesn't really change the reference models, so we can move forward without them. 
-   All modified models w/o a temporal resolution or depth layer were all within 0.05 TSS and AUC.
-   The combined AGI and DO model performed poorly. Will be best to keep them separated. 

