---
title: "Mako hSDM covariate explore (ROMS and CMEM domains)"
author: "Emily Nazario"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
---

```{r}
#| include: false
library(tidyverse)
library(here)
library(terra)
library(tidyquant)
library(MetBrewer)
library(tidyquant)

cmem_dat <- readRDS(here("data/locs_w_covar/cmems/cmem_locs_covar_AGI_0m.rds"))
cmem_dat <- cmem_dat %>% rename("pO2_0_atm" = "pO2_0") #removes oxy (atm) col. Was needed to calc AGI

theme_set(theme_tq() +
  theme(strip.text = element_text(size = 16), 
        axis.text = element_text(size = 14, color = "black"), 
        axis.title = element_text(size = 16), 
        legend.text = element_text(size = 16)))

```

For each of the figures below, I have visualized what the covariate values are for the CRW PA positions and observed positions. Some values were extracted using the ROMS data, and thus limiting the position range to this domain. I have also explored the covariate values using the full range of the available positions using covariate files from Copernicus (CMEM). Please note that the lists of covariates selected for each data source do vary depending on what was recommended and available.

The covariate values explored on this page are all from the surface, and thus depth is not yet considered in this analysis.

Lastly, for each collection of a shark's observed positions, there are 32 replicate PA CRW tracks for the CMEM domain and 16 replicate tracks for the ROMS domain.

# Covariate summary plots for all sharks

## CMEM figures

::: panel-tabset
### Map of obsv. positions
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 9

#define map for figure
north_map = map_data("world") %>% group_by(group)
shore     = north_map[north_map$region=="Canada" 
                      | north_map$region=="USA"
                      | north_map$region=="Mexico",]

cmem_dat2 <- cmem_dat %>% filter(PA == 0)

ggplot(shore, aes(long, lat)) +
  coord_map("mercator", xlim = c(-140, -110), ylim = c(10, 50))+
  geom_point(data=cmem_dat2, aes(lon, lat, colour=as.factor(tag))) +
  geom_polygon(aes(group=group), fill="grey60",lwd=1) +
  scale_color_manual(values = met.brewer("Hokusai2", 23))+
  theme(legend.position = "right", 
        legend.key.size = unit(0.75,'cm'), 
        legend.text = element_text(size = 10), 
        panel.spacing.x = unit(0, "lines"),
        panel.spacing.y = unit(0, "lines"),
        legend.margin=margin(0,0,0,0))

```

### Density

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 12

cmem_0m_long <- gather(cmem_dat, covar, value, thetao_mean:AGI_0m) %>% mutate(PA = as.factor(PA))

ggplot(cmem_0m_long, aes(x = value, fill = PA)) + 
  geom_density(alpha = 0.35) + 
  facet_wrap(~covar, scales = "free") + 
  scale_fill_manual(values = c("#ABC9C8", "#0E3F61"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Boxplots

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 12

ggplot(cmem_0m_long, aes(x = PA, y = value, fill = PA)) + 
  geom_boxplot() + 
  facet_wrap(~covar, scales = "free")+ 
  scale_fill_manual(values = c("#ABC9C8", "#0E3F61"))
```

### Time vs. covar scatter

The figures below only represent covariate values for observed positions. PA positions were omitted from this analysis due to the quantity.

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 12
#| fig-height: 10

cmem_0m_long %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = dt, y = value, color = covar)) + 
  geom_point(alpha = 0.2) +
  geom_smooth(color = "grey20", se = FALSE)+
  facet_wrap(~covar, scales = "free") + 
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 13))+
  xlab("")
```
:::

### Covar vs. covar scatter

Below are some comparisons of the covariate values relative to each other for the observed positions (covariate values at PA positions are excluded).

::: panel-tabset
#### DO (atm) vs. AGI

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = pO2_0_atm, y = AGI_0m, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Temp vs. AGI

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = thetao_mean, y = AGI_0m, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Temp vs. DO (atm)

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = thetao_mean, y = pO2_0_atm, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### DO (mmol/m^3) vs. MLD

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = o2_mean, y = mlotst_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Lat vs. AGI

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lat, y = AGI_0m, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Lat vs. DO (mmol/m^3)

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lat, y = o2_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Lon vs. AGI

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lon, y = AGI_0m, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Lon vs. DO (mmol/m^3)

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lon, y = o2_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Bathy vs. AGI

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = bathy, y = AGI_0m, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Bathy vs. DO (mmol/m^3)

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = bathy, y = o2_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Lat vs. Bathy

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lat, y = bathy, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Lon vs. Bathy

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lon, y = bathy, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Lat vs. Temp

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lat, y = thetao_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Lat vs. Temp vs. DO (mmol/m^3)

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 6
#| fig-height: 6
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lat, y = thetao_mean, color = o2_mean)) + 
  geom_point(alpha = 0.3)+
  theme(legend.position = "right")
```

#### Lon vs. Temp vs. DO (mmol/m^3)

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 6
#| fig-height: 6
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lon, y = thetao_mean, color = o2_mean)) + 
  geom_point(alpha = 0.3)+
  theme(legend.position = "right")
```

#### Temp vs. MLD

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = thetao_mean, y = mlotst_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```

#### Lat vs. MLD

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4
#| fig-align: center

cmem_dat %>% 
  filter(PA == 0) %>%
  ggplot(aes(x = lat, y = mlotst_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.2) +
  scale_color_manual(values = c("#0a3351"))+
  theme(legend.position = "none")
```
:::

## ROMS figures

::: panel-tabset
### Map of obsv. positions

### Density

### Boxplots

### Time vs. covar value scatter
:::

### Covar vs. covar scatter

::: panel-tabset
#### Temp vs. DO

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 4
#| fig-height: 4

#null 4 now
```
:::

## Select shark analyses
I randomly selected three tag IDs to include for this next analysis.
```{r}
set.seed(1004)
sample(1:23, 3) #randomly select 3 sharks: 11, 22, and 1

#filter to only keep the tag IDs of the above three sharks.
cmem_dat_sub <- cmem_dat %>% filter(tag == "41770" | tag == "68285" | tag == "78151")
cmem_obsv_sub <- cmem_dat_sub %>% filter(PA == 0)
```

I then randomly select which repetition number I want to include as a representative for this series of analyses. I am left with a subset of the data which will be used for the Select Shark Analyses portion of this document.
```{r}
sample(1:32, 1) 
cmem_rep_sub = NULL
for(i in 1:length(unique(cmem_dat_sub$tag))){
  curr_ID <- unique(cmem_dat_sub$tag)[i]
  temp_df <- cmem_dat_sub[cmem_dat_sub$tag %in% curr_ID,]
  temp_df <- temp_df %>% group_by(rep)
  
  curr_rep <- unique(temp_df$rep)[13] #number randomly selected above
  temp_df2 <- temp_df[temp_df$rep %in% curr_rep,]
  
  cmem_rep_sub <- rbind(cmem_rep_sub, temp_df2)
}

cmem_dat_sub2 <- rbind(cmem_obsv_sub, cmem_rep_sub)
```

### CMEM AGI
These figures track how DO changes through a shark's track for both its observed positions ('0') and PA positions ('1').

::: panel-tabset
#### track values vs. space/time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 10

ggplot(shore, aes(long, lat)) +
  coord_map("mercator", xlim = c(-140, -110), ylim = c(10, 50))+
  geom_point(data = cmem_dat_sub2, aes(lon, lat, colour = AGI_0m), size = 2) +
  geom_polygon(aes(group = group), fill = "grey60",lwd = 1) +
  facet_grid(tag~PA)+
  theme(legend.position = "right", 
        legend.key.size = unit(0.75,'cm'), 
        legend.text = element_text(size = 10), 
        panel.spacing.x = unit(0, "lines"),
        panel.spacing.y = unit(0, "lines"),
        legend.margin=margin(0,0,0,0))


```

#### track values vs. time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(cmem_dat_sub2, aes(x = dt, y = AGI_0m, color = as.factor(PA))) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE)+
  facet_grid(~tag, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 2))+
  xlab("")

```

#### track values vs. lat
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(cmem_dat_sub2, aes(x = lat, y = AGI_0m, color = as.factor(PA))) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE)+
  facet_grid(~tag, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 2))

```

#### track values vs. lon
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(cmem_dat_sub2, aes(x = lon, y = AGI_0m, color = as.factor(PA))) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE)+
  facet_grid(~tag, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 2))

```
:::

### CMEM DO
These figures track how DO changes through a shark's track for both its observed positions ('0') and PA positions ('1').

::: panel-tabset
#### track values vs. space/time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 10

ggplot(shore, aes(long, lat)) +
  coord_map("mercator", xlim = c(-140, -110), ylim = c(10, 50))+
  geom_point(data = cmem_dat_sub2, aes(lon, lat, colour = o2_mean), size = 2) +
  geom_polygon(aes(group = group), fill = "grey60",lwd = 1) +
  facet_grid(tag~PA)+
  theme(legend.position = "right", 
        legend.key.size = unit(0.75,'cm'), 
        legend.text = element_text(size = 10), 
        panel.spacing.x = unit(0, "lines"),
        panel.spacing.y = unit(0, "lines"),
        legend.margin=margin(0,0,0,0))


```

#### track values vs. time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(cmem_dat_sub2, aes(x = dt, y = o2_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE)+
  facet_grid(~tag, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 2))+
  xlab("")

```

#### track values vs. lat
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(cmem_dat_sub2, aes(x = lat, y = o2_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE)+
  facet_grid(~tag, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 2))

```

#### track values vs. lon
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(cmem_dat_sub2, aes(x = lon, y = o2_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE)+
  facet_grid(~tag, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 2))

```
:::

### CMEM Temp.
These figures track how temp. changes through a shark's track for both its observed positions ('0') and PA positions ('1').

::: panel-tabset
#### track values vs. space/time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 10

ggplot(shore, aes(long, lat)) +
  coord_map("mercator", xlim = c(-140, -110), ylim = c(10, 50))+
  geom_point(data = cmem_dat_sub2, aes(lon, lat, colour = thetao_mean), size = 2) +
  geom_polygon(aes(group = group), fill = "grey60",lwd = 1) +
  facet_grid(tag~PA)+
  theme(legend.position = "right", 
        legend.key.size = unit(0.75,'cm'), 
        legend.text = element_text(size = 10), 
        panel.spacing.x = unit(0, "lines"),
        panel.spacing.y = unit(0, "lines"),
        legend.margin=margin(0,0,0,0))


```

#### track values vs. time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(cmem_dat_sub2, aes(x = dt, y = thetao_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE)+
  facet_grid(~tag, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 2))+
  xlab("")

```

#### track values vs. lat
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(cmem_dat_sub2, aes(x = lat, y = thetao_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE)+
  facet_grid(~tag, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 2))

```

#### track values vs. lon
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

ggplot(cmem_dat_sub2, aes(x = lon, y = thetao_mean, color = as.factor(PA))) + 
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE)+
  facet_grid(~tag, scales = "free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_color_manual(values = met.brewer("Hokusai2", 2))

```
:::

### ROMS DO
These figures track how DO changes through a shark's track for both its observed positions ('0') and PA positions ('1').

::: panel-tabset
#### track values vs. space/time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 10



```

#### track values vs. time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6



```

#### track values vs. lat
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6



```

#### track values vs. lon
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6



```
:::

### ROMS Temp.
These figures track how temp. changes through a shark's track for both its observed positions ('0') and PA positions ('1').

::: panel-tabset
#### track values vs. space/time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 10



```

#### track values vs. time
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6



```

#### track values vs. lat
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6



```

#### track values vs. lon
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6



```
:::

## Maps

Plot DO and temp. values over space for a few representative dates for sanity check of mapped values. Randomly selected three days (day)

### CMEM

::: panel-tabset

#### DO (mmol/L)
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6
set.seed(1004)
#sample(1:2192, 3) #gives the date layers selected below
cmem_rast <- rast(here("data/enviro/CMEMS/processed/CMEM_SST_SAL_MLD_SSH_UO_VO_CHL_NPP_DO_0m_Jan2004_Dec2009_0.25_D.nc"))

plot(cmem_rast$o2_1419)
plot(cmem_rast$o2_1878)
plot(cmem_rast$o2_52)
```

#### Temp.(*C)
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

plot(cmem_rast$thetao_1419)
plot(cmem_rast$thetao_1878)
plot(cmem_rast$thetao_52)

```

#### Bathy (m)
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6

bathy_rast <- rast(here("data/enviro/bathy/processed/gebco_bathy_cmems_domain.nc"))
plot(bathy_rast)
```
:::

### ROMS

::: panel-tabset
#### DO
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6


```

#### Temp.
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 6



```
:::
