---
title: "Diet exploration"
author: "Emily Nazario"
date: "`r Sys.Date()`"
format:
 html: 
  embed-resources: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
---

```{r}
#| echo: false
#| message: false
#| label: Load libraries
#| include: false

library(tidyverse)
library(tidyquant)
library(kableExtra)
library(here)

options(scipen=999)
```

```{r}
#| echo: false
#| message: false
#| label: Import and format data
#| include: false

dat_coll <- read.csv(here("data/diet/collection.csv"))
dat_pred <- read.csv(here("data/diet/predator.csv"))
dat_prey <- read.csv(here("data/diet/prey_comp.csv"))
dat_prey_size <- read.csv(here("data/diet/prey_size.csv"))

  #limits diet data to SCB Ecoregion
scbE_coll <- dat_coll %>% 
  filter(Latitude < 47.33 & Latitude > 2.76 & Longitude > -150.94 & Longitude < -105.63)

#filters predator DF based on IDs from SCB Ecoregion described above. Filters out FL that weren't in the loc DF
dat_pred <- dat_pred %>% #filters out measurement entries that are blank
  filter(Predator_Measurement_2 == "FL") %>%
  filter()

scbE_coll_id <- scbE_coll$Collection_ID #438 IDs
scbE_pred <- dat_pred[dat_pred$Collection_ID %in% scbE_coll_id, ]
FL_femaleD <- scbE_pred %>%
  filter(Predator_Sex == "F") %>%
  filter(Predator_Length_2 < 249 & Predator_Length_2 > 100)

FL_maleD <- scbE_pred %>%
  filter(Predator_Sex == "M") %>%
  filter(Predator_Length_2 < 180 & Predator_Length_2 > 100)

scbE_pred2 <- rbind(FL_femaleD, FL_maleD)

#filters prey DF based on IDs from SCB and FL filtered predator DF 
scbE_pred_id <- scbE_pred2$Predator_ID #117 IDs
scbE_prey <- dat_prey[dat_prey$Predator_ID %in% scbE_pred_id, ]

#### RMPQ and GII Calculations -- all years#####
scbE_prey <- scbE_prey %>%
  mutate(freq_metric = 1)

rmpq_prey <- scbE_prey %>%
  group_by(Common_Name) %>%
  summarise(N = sum(Prey_N), 
            W = sum(Prey_Wt), 
            Freq = sum(freq_metric), 
            perc_N = (N/1353)*100, 
            perc_W = (W/90099.18)*100, 
            perc_F = (Freq/264)*100, 
            GII = (perc_N + perc_W + perc_F)/sqrt(3), 
            perc_GII = (perc_N + perc_W + perc_F)/3)
```

## Diet data by year

Here is the diet data by year for the area, FL range, and time period included in the new study parameters. All columns should be how they are typically referenced with the addition of the `pred_metric` column. This is a new one I added representing the unique number of stomachs for that year, which was used to calculate the percent frequency value for each prey item.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| label: Data preview
#Match dfs by IDs and bring over the collection year to the prey DF
yr_coll <- scbE_coll %>%
  subset(select = c("Collection_ID", "Year"))

scbE_pred3 <- merge(scbE_pred2, yr_coll, by = "Collection_ID", all = T)
yr_pred <- scbE_pred3 %>%
  subset(select = c("Year", "Predator_ID"))

scbE_prey2 <- merge(scbE_prey, yr_pred, by = "Predator_ID", all = T) %>%
  drop_na(Prey_ID)

#calculate the RMPQ values by species and year
rmpq_prey_year <- scbE_prey2 %>%
  group_by(Year) %>%
  mutate(pred_metric = length(unique(Predator_ID))) %>%
  ungroup() %>%
  group_by(Year, Common_Name) %>%
  summarise(N = sum(Prey_N), 
            W = sum(Prey_Wt), 
            Freq = sum(freq_metric), 
            pred_metric = unique(pred_metric)) %>%
  ungroup() %>%
  filter(Year != 1998)

#calculate the GII by year using the by species data
rmpq_prey_year2 <- rmpq_prey_year %>%
  group_by(Year) %>%
  mutate(perc_N = (N/sum(N))*100, 
         perc_W = (W/sum(W))*100, 
         perc_F = (Freq/pred_metric)*100,
         GII = (perc_N + perc_W + perc_F)/sqrt(3), 
         perc_GII = (perc_N + perc_W + perc_F)/3) %>%
  filter(Year <= 2013 & Year >= 2003)

#generate the formatted quarto table 
kbl(rmpq_prey_year2) %>% kable_paper("striped") %>% scroll_box(width = "100%", height = "400px")
```

## GII bar plots

Below are bar plots of the %GII for each prey species by year. Plots are organized from highest to lowest %GII. These %GII values were calculated using the data table above (filtered according to the new study parameters). %GII values less than 1 were not plotted. An important note is that weak El Ni単o events occurred in 2004-2005 and 2006-2007, while moderate El Ni単o events occurred in 2009-2010. Additionally, weak La Ni単a events occurred in 2005-2006 and 2008-2009, while strong La Ni単a events occurred in 2007-2008.

::: panel-tabset
## All years

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: All years 
rmpq_prey %>%
  filter(perc_GII >= 1) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')
```

## 2003

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand three
rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2003) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2004

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand four
rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2004) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2005

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand five

rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2005) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2006

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand six

rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2006) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2007

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand seven

rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2007) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2008

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand eight

rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2008) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2009

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand nine

rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2009) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2010

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand ten

rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2010) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2011

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand eleven

rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2011) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2012

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand twelve

rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2012) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```

## 2013

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}
#| label: two thousand thirteen

rmpq_prey_year2 %>%
  filter(perc_GII >= 1 & Year == 2013) %>% 
  ggplot(aes(x = reorder(Common_Name, -perc_GII), y = perc_GII)) +
  geom_bar(stat = "identity") +
  theme_tq() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 0.95), 
        axis.text=element_text(size=14), 
        axis.title=element_text(size=14)) +
  xlab('')+
  ylab('% GII')

```
:::
