---
title: "Mercator dissolved oxygen and temperature validation"
author: "Emily Nazario"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
code-block-bg: true
---

On this document I compare the mean Mercator BIORYS4 and GLORYS4 dissolved oxygen (DO) and temperature (temp) data to values from climatology products (DO: World Ocean Atlas; temp: OISST and WOA) and bottle collections at CalCOFI stations. DO and temp values will be compared to the climatology products as averages across the study area (lat: 1 to 49 and lon: -153 to -104) at 0m and at 200m. While the comparisons to the CalCofi data will be done at the nearest grid cell for each station both at 0m and 20m. These spatial averages will also be compared on annual, seasonal, and monthly scales.

Temporal variability was compared by extracting the mean dissolved oxygen and temperature values over the CalCOFI domain for each day a CalCOFI sample was performed throughout the study period (Jun 26th, 2003- Mar. 22nd, 2015). I will then compare the CalCOFI and Mercator dissolved oxygen and temperature means over time.

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: libraries

### libraries #### 
library(tidyverse)
library(terra)
library(sf)
library(here)
library(ncdf4)
library(respR)
library(patchwork)
library(kableExtra)
library(tidyterra)
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Mercator data
## data ####
merc_do0_raw <- rast(here("data/enviro/model_validate/Mercator/CMEMS_CHL_DO_0m_JAN2003_Dec2015_0.25deg_D.nc"))
merc_do200_raw <- rast(here("data/enviro/model_validate/Mercator/CMEMS_CHL_DO_200m_JAN2003_Dec2015_0.25deg_D.nc"))
merc_temp0_raw <- list.files(here("data/enviro/model_validate/Mercator"), pattern = "Temp_0m", full.names = TRUE)
merc_temp200_raw <- list.files(here("data/enviro/model_validate/Mercator"), pattern = "Temp_200m", full.names = TRUE)

clim_do_ann <- rast(here("data/enviro/model_validate/Climatology/woa_oxygen/woa18_all_o00_01.nc"))
clim_do_mo <- list.files(here("data/enviro/model_validate/Climatology/woa_oxygen/monthly"), full.names = TRUE)
clim_tempWA <- rast(here("data/enviro/model_validate/Climatology/sst.mon.mean.nc"))

calcofi_samp <- read.csv(here("data/enviro/model_validate/CalCofi_IMECOCAL/CAlCOFI/194903-202105_Bottle.csv"), check.names = FALSE)
calcofi_met <- read.csv(here("data/enviro/model_validate/CalCofi_IMECOCAL/CAlCOFI/194903-202105_Cast.csv"), check.names = FALSE)
#imecocal

### Mercator data ####
#### oxygen (input: mmol/m^3, now: mmol/L) #### 
##### 0m #####
merc_do0_raw

#filter to separate variables by the "=" and prep to be able to filter by variable in next step
merc_vars0 <- str_split(names(merc_do0_raw), "=") %>% map_chr(\(x) x[1]) %>% str_extract("[^_]+")

#identify location of oxygen layers
merc_pos0 <- which(merc_vars0 == "o2")

#subset to only keep oxygen layers
merc_do0 <- merc_do0_raw %>% subset(merc_pos0) 
merc_do0 #spatial res: 0.25*, temporal es: daily, crs: lon/lat WGS 84

#convert DO to mmol/L
#convert temporal resolution to monthly and annually. Convert DO values from mmol/m^3 to mmol/L.
merc_do_M0 <- tapp(merc_do0, "months", mean)
merc_do_M0 <- merc_do_M0/1000
merc_do_Y0 <- tapp(merc_do0, "years", mean)
merc_do_Y0 <- merc_do_Y0/1000

#convert monthly resolution to seasonal 
#winter
# merc_do_Wn0 <- subset(merc_do_M0, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
# merc_do_Wn0 <- app(merc_do_Wn0, mean)
# #spring
# merc_do_Sp0 <- subset(merc_do_M0, c("m_3", "m_4", "m_5")) #Mar/Apr/May
# merc_do_Sp0 <- app(merc_do_Sp0, mean)
# #summer
# merc_do_Su0 <- subset(merc_do_M0, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
# merc_do_Su0 <- app(merc_do_Su0, mean)
# #fall
# merc_do_Fa0 <- subset(merc_do_M0, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
# merc_do_Fa0 <- app(merc_do_Fa0, mean)

##### 200m #####
#filter to separate variables by the "=" and prep to be able to filter by variable in next step
merc_vars200 <- str_split(names(merc_do200_raw), "=") %>% map_chr(\(x) x[1]) %>% str_extract("[^_]+")

#identify location of oxygen layers
merc_pos200 <- which(merc_vars200 == "o2")

#subset to only keep oxygen layers
merc_do200 <- merc_do200_raw %>% subset(merc_pos200) 
merc_do200 #spatial res: 0.25*, temporal es: daily, crs: lon/lat WGS 84

#convert DO to mmol/L
#convert temporal resolution to monthly and annually. Convert DO from mmol/m^3 to mmol/L.
merc_do_M200 <- tapp(merc_do200, "months", mean)
merc_do_M200 <- merc_do_M200/1000
merc_do_Y200 <- tapp(merc_do200, "years", mean)
merc_do_Y200 <- merc_do_Y200/1000

# #convert monthly resolution to seasonal 
# #winter
# merc_do_Wn200 <- subset(merc_do_M200, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
# merc_do_Wn200 <- app(merc_do_Wn200, mean)
# #spring
# merc_do_Sp200 <- subset(merc_do_M200, c("m_3", "m_4", "m_5")) #Mar/Apr/May
# merc_do_Sp200 <- app(merc_do_Sp200, mean)
# #summer
# merc_do_Su200 <- subset(merc_do_M200, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
# merc_do_Su200 <- app(merc_do_Su200, mean)
# #fall
# merc_do_Fa200 <- subset(merc_do_M200, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
# merc_do_Fa200 <- app(merc_do_Fa200, mean)

#### temperature (C) #### 
##### 0m #####
#combine all temp rasts into two rasters separated by depth layer

merc_temp_all0 <- NULL
for(i in 1:length(merc_temp0_raw)){
  # temp_nc <- nc_open(merc_temp0_raw[[i]])
  # 
  # #get correct lat and lon info from ncdf4 package
  # mod_rast <- rast(
  #   rast_lat <- ncvar_get(temp_nc, "nav_lat"),
  #   rast_lon <- ncvar_get(temp_nc, "nav_lon")
  # )
  # 
  # temp_lat <- rast_lat[1,1:241]
  # temp_lon <- rast_lon[1,1:241]
  
  temp_rast <- rast(merc_temp0_raw[[i]])
  rast_source <- sources(temp_rast)
  rast_year <- str_extract_all(rast_source, "\\d{4}")[[1]][1]
  
  d1 <- seq(as.Date(paste0(rast_year,"-","01-01")), by = "month",length = 6)
  d2 <- seq(as.Date(paste0(rast_year,"-","07-01")), by = "month",length = 6)
  
  #add year-month info to raster files
  if(str_detect(rast_source, "JAN") == TRUE){
    time(temp_rast) <- d1
  } else {
    time(temp_rast) <- d2
  }
  
  merc_temp_all0 <- append(merc_temp_all0, temp_rast)
}

merc_temp_all0 #spatial res: 0.25*, temporal es: year-month, crs: lon/lat WGS 84

#convert temporal resolution to monthly and annually
merc_temp_M0 <- tapp(merc_temp_all0, "months", mean)
merc_temp_Y0 <- tapp(merc_temp_all0, "years", mean)

#convert monthly resolution to seasonal 
#winter
# merc_temp_Wn0 <- subset(merc_temp_M0, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
# merc_temp_Wn0 <- app(merc_temp_Wn0, mean)
# #spring
# merc_temp_Sp0 <- subset(merc_temp_M0, c("m_3", "m_4", "m_5")) #Mar/Apr/May
# merc_temp_Sp0 <- app(merc_temp_Sp0, mean)
# #summer
# merc_temp_Su0 <- subset(merc_temp_M0, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
# merc_temp_Su0 <- app(merc_temp_Su0, mean)
# #fall
# merc_temp_Fa0 <- subset(merc_temp_M0, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
# merc_temp_Fa0 <- app(merc_temp_Fa0, mean)

##### 200m #####
#combine all temp rasts into two rasters separated by depth layer
merc_temp_all200 <- NULL
for(i in 1:length(merc_temp200_raw)){
  temp_rast <- rast(merc_temp200_raw[[i]])
  rast_source <- sources(temp_rast)
  rast_year <- str_extract_all(rast_source, "\\d{4}")[[1]][1]
  
  d1 <- seq(as.Date(paste0(rast_year,"-","01-01")), by = "month",length = 6)
  d2 <- seq(as.Date(paste0(rast_year,"-","07-01")), by = "month",length = 6)
  
  #add year-month info to raster files
  if(str_detect(rast_source, "JAN") == TRUE){
    time(temp_rast) <- d1
  } else {
    time(temp_rast) <- d2
  }
  
  merc_temp_all200 <- append(merc_temp_all200, temp_rast)
}

merc_temp_all200 #spatial res: 0.25*, temporal es: year-month, crs: lon/lat WGS 84

#convert temporal resolution to monthly and annually
merc_temp_M200 <- tapp(merc_temp_all200, "months", mean)
merc_temp_Y200 <- tapp(merc_temp_all200, "years", mean)

#convert monthly resolution to seasonal 
# #winter
# merc_temp_Wn200 <- subset(merc_temp_M200, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
# merc_temp_Wn200 <- app(merc_temp_Wn200, mean)
# #spring
# merc_temp_Sp200 <- subset(merc_temp_M200, c("m_3", "m_4", "m_5")) #Mar/Apr/May
# merc_temp_Sp200 <- app(merc_temp_Sp200, mean)
# #summer
# merc_temp_Su200 <- subset(merc_temp_M200, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
# merc_temp_Su200 <- app(merc_temp_Su200, mean)
# #fall
# merc_temp_Fa200 <- subset(merc_temp_M200, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
# merc_temp_Fa200 <- app(merc_temp_Fa200, mean)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Climatology data

### Climatology data ####
#### oxygen (input: umol/kg, now: mmol/L) ####
#define extent to study area
e <- ext(-153, -104, 1, 49)
##### 0m #####
#decadal values for o2 conversion at surface for annual average
temp_o20 <- rast(here("data/enviro/model_validate/Climatology/woa_oxygen/temp_for_conversions/woa18_decav_t00_01.nc"))
temp_o20 <- crop(temp_o20,e)
temp_o20 <- temp_o20$`t_an_depth=0`

sal_o20 <- rast(here("data/enviro/model_validate/Climatology/woa_oxygen/sal_for_conversions/woa18_decav_s00_01.nc"))
sal_o20 <- crop(sal_o20,e)
sal_o20 <- sal_o20$`s_an_depth=0`

#spatial res: 0.1*, temporal res: averaged annually, temporal range: 1960-2017
#annual
clim_do_ann0W <- clim_do_ann$`o_mn_depth=0` 

clim_do_ann0 <- crop(clim_do_ann0W, e)
#plot(clim_do_ann0)

#convert to mmol/L
clim_doY_mmol_L <- clim_do_ann0*1.025*0.001 #1.025 kg/L and 0.001mmol in umol
#plot(clim_doY_mmol_L)

#monthly and seasonal files
clim_do_mo0 = NULL
for(i in 1:length(clim_do_mo)){
  temp_rast <- rast(clim_do_mo[[i]])
  temp_rast_mean <- temp_rast$`o_mn_depth=0`
  temp_rast_mean <- crop(temp_rast_mean, e)
  
  rast_name <- paste0("m", "_", i)
  names(temp_rast_mean) <- rast_name
  
  clim_do_mo0 <- append(clim_do_mo0, temp_rast_mean)
}

#monthly
clim_doM_mmol_L <- clim_do_mo0*1.025*0.001

# #convert monthly resolution to seasonal 
# #winter
# clim_do_Wn0 <- subset(clim_doM_mmol_L, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
# clim_do_Wn0 <- app(clim_do_Wn0, mean)
# #spring
# clim_do_Sp0 <- subset(clim_doM_mmol_L, c("m_3", "m_4", "m_5")) #Mar/Apr/May
# clim_do_Sp0 <- app(clim_do_Sp0, mean)
# #summer
# clim_do_Su0 <- subset(clim_doM_mmol_L, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
# clim_do_Su0 <- app(clim_do_Su0, mean)
# #fall
# clim_do_Fa0 <- subset(clim_doM_mmol_L, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
# clim_do_Fa0 <- app(clim_do_Fa0, mean)

##### 200m #####
#spatial res: 0.1*, temporal res: averaged annually, temporal range: 1960-2017
clim_do_ann200W <- clim_do_ann$`o_mn_depth=200` 
clim_do_ann200 <- crop(clim_do_ann200W, e)

clim_doY_mmol_L200 <- clim_do_ann200*1.025*0.001
#plot(clim_doY_mmol_L200)

#need to do monthly and seasonal files
clim_do_mo200 = NULL
for(i in 1:length(clim_do_mo)){
  temp_rast <- rast(clim_do_mo[[i]])
  temp_rast_mean <- temp_rast$`o_mn_depth=200`
  temp_rast_mean <- crop(temp_rast_mean, e)
  
  rast_name <- paste0("m", "_", i)
  names(temp_rast_mean) <- rast_name
  
  clim_do_mo200 <- append(clim_do_mo200, temp_rast_mean)
}

clim_doM_mmol_L200 <- clim_do_mo200*1.025*0.001

# #convert monthly resolution to seasonal 
# #winter
# clim_do_Wn200 <- subset(clim_doM_mmol_L200, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
# clim_do_Wn200 <- app(clim_do_Wn200, mean)
# global(clim_do_Wn200, "mean", na.rm=TRUE)
# 
# #spring
# clim_do_Sp200 <- subset(clim_doM_mmol_L200, c("m_3", "m_4", "m_5")) #Mar/Apr/May
# clim_do_Sp200 <- app(clim_do_Sp200, mean)
# global(clim_do_Sp200, "mean", na.rm=TRUE)
# 
# #summer
# clim_do_Su200 <- subset(clim_doM_mmol_L200, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
# clim_do_Su200 <- app(clim_do_Su200, mean)
# global(clim_do_Su200, "mean", na.rm=TRUE)
# 
# #fall
# clim_do_Fa200 <- subset(clim_doM_mmol_L200, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
# clim_do_Fa200 <- app(clim_do_Fa200, mean)
# global(clim_do_Fa200, "mean", na.rm=TRUE)

#### temperature (C) ####
##### 0m #####
#clim_tempWA #spatial res: 0.25*, temporal res: annual, temporal range: 1981-2024
#plot(clim_tempWA)

e2 <- ext(207, 256, 1, 49)
clim_tempA <- crop(clim_tempWA, e2)
#plot(clim_tempA$sst_1)

clim_tempA <- subset(clim_tempA, time(clim_tempA) > as.Date("2003-01-01"))
clim_tempA <- subset(clim_tempA, time(clim_tempA) < as.Date("2015-12-31"))
#clim_tempA

#convert temporal resolution to monthly and annually
clim_temp_M0 <- tapp(clim_tempA, "months", mean)
clim_temp_Y0 <- tapp(clim_tempA, "years", mean)

#convert monthly resolution to seasonal 
# #winter
# clim_temp_Wn0 <- subset(clim_temp_M0, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
# clim_temp_Wn0 <- app(clim_temp_Wn0, mean)
# #spring
# clim_temp_Sp0 <- subset(clim_temp_M0, c("m_3", "m_4", "m_5")) #Mar/Apr/May
# clim_temp_Sp0 <- app(clim_temp_Sp0, mean)
# #summer
# clim_temp_Su0 <- subset(clim_temp_M0, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
# clim_temp_Su0 <- app(clim_temp_Su0, mean)
# #fall
# clim_temp_Fa0 <- subset(clim_temp_M0, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
# clim_temp_Fa0 <- app(clim_temp_Fa0, mean)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Station data

### Station data ####
#### CalCOFI ####
#filter to correct dates
calcofi_met_filt <- calcofi_met %>%
  filter(Year >= 2003 & Year <= 2015)

calcofi_samp_filt <- calcofi_samp[calcofi_samp$Cst_Cnt %in% calcofi_met_filt$Cst_Cnt, ]

#subset metadata to just Cst_Cnt ID, year, and month to join to sample DF
calcofi_met_sub <- calcofi_met_filt %>% 
  subset(select = c(Cst_Cnt, Year, Month))

calcofi_samp_filt1 <- calcofi_samp_filt %>%
  full_join(calcofi_met_sub, by = "Cst_Cnt")

calcofi_samp_filt1 <- calcofi_samp_filt1 %>%
  mutate(season = ifelse(Month >= 3 & Month <= 5, "Spring", 
                         ifelse(Month >= 6 & Month <= 8, "Summer", 
                                ifelse(Month >= 9 & Month <= 11, "Fall", "Winter"))))

##### 0m #####
calcofi_samp0 <- calcofi_samp_filt1 %>% 
  filter(Depthm == 0) %>%
  filter(!is.na(O2ml_L)) %>%
  mutate(do0_mmol_L = O2ml_L/22.3916)

#convert DO to mmol/L for each row either using the average temp/sal if data isn't available, or the in situ measurements at that station/bottle collection
# mean_sal = mean(calcofi_samp0$Salnty, na.rm = TRUE)
# mean_temp = mean(calcofi_samp0$T_degC, na.rm = TRUE)
# calcofi_samp0$do0_mmol_L = NA
# 
# for(i in 1:nrow(calcofi_samp0)){
#   if(is.na(calcofi_samp0$Salnty[i]) | is.na(calcofi_samp0$T_degC[i])){
#     
#    calcofi_samp0$do0_mmol_L[i] = convert_DO(calcofi_samp0$O2ml_L[i], from = "mL/L", to = "mmol/L", S = mean_sal, t = mean_temp, P = 1.013253)
#    
#   } else if(!is.na(calcofi_samp0$Salnty[i]) & !is.na(calcofi_samp0$T_degC[i])){
#     
#     calcofi_samp0$do0_mmol_L[i] = convert_DO(calcofi_samp0$O2ml_L[i], from = "mL/L", to = "mmol/L", S = calcofi_samp0$Salnty[i], t = calcofi_samp0$T_degC[i], P = 1.013253)
#     
#   }
#   #print(calcofi_samp0$do0_mmol_L[i])
# }

##### 200m #####
calcofi_samp200 <- calcofi_samp_filt1 %>%
  filter(Depthm == 200) %>%
  filter(!is.na(O2ml_L))%>%
  mutate(do200_mmol_L = O2ml_L/22.3916)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Modifying merc and clim domains to calcofi

#subset metadata to just Cst_Cnt ID, year, and month to join to sample DF
calcofi_coords <- calcofi_met_filt %>% 
  subset(select = c(Cst_Cnt, Lat_Deg, Lon_Deg))

calcofi_samp_coords <- calcofi_samp0 %>%
  full_join(calcofi_coords, by = "Cst_Cnt")

calcofi_samp_coords200 <- calcofi_samp200 %>%
  full_join(calcofi_coords, by = "Cst_Cnt")
# min(calcofi_samp_coords$Lat_Deg)
# max(calcofi_samp_coords$Lat_Deg)
# min(calcofi_samp_coords$Lon_Deg)
# max(calcofi_samp_coords$Lon_Deg)

e_cal <- ext(-125, -117, 29, 37)
e_cal_m <- rast(xmin=650, xmax=680, ymin=621, ymax=660, nrows=25, ncols=25)
e_cal_cl <- ext(235, 243, 29, 37)

#Merc 0m - Y/S/M
#annual 
#do
merc_do_Y0_cal <- crop(merc_do_Y0, e_cal)

#temp
merc_temp_Y0_cal <- crop(merc_temp_Y0, e_cal_m)

#seasonal 
#do
# merc_do_Wn0_cal <- crop(merc_do_Wn0, e_cal)
# merc_do_Sp0_cal <- crop(merc_do_Sp0, e_cal)
# merc_do_Su0_cal <- crop(merc_do_Su0, e_cal)
# merc_do_Fa0_cal <- crop(merc_do_Fa0, e_cal)
# 
# #temp
# merc_temp_Wn0_cal <- crop(merc_temp_Wn0, e_cal_m)
# merc_temp_Sp0_cal <- crop(merc_temp_Sp0, e_cal_m)
# merc_temp_Su0_cal <- crop(merc_temp_Su0, e_cal_m)
# merc_temp_Fa0_cal <- crop(merc_temp_Fa0, e_cal_m)
# 
# #month
# #do
# merc_do_M0_cal <- crop(merc_do_M0, e_cal)
# 
# #temp
# merc_temp_M0_cal <- crop(merc_temp_M0, e_cal_m)
# 
# #Merc 200m - Y/S/M -- DO PENDING ON RESULTS
# #annual
# #temp 
# merc_temp_Y200_cal <- crop(merc_temp_Y200, e_cal_m)
# 
# #seasonal
# #temp
# merc_temp_Wn200_cal <- crop(merc_temp_Wn200, e_cal_m)
# merc_temp_Sp200_cal <- crop(merc_temp_Sp200, e_cal_m)
# merc_temp_Su200_cal <- crop(merc_temp_Su200, e_cal_m)
# merc_temp_Fa200_cal <- crop(merc_temp_Fa200, e_cal_m)
# 
# #month
# #temp
# merc_temp_M200_cal <- crop(merc_temp_M200, e_cal_m)
# 
# 
# #Clim 0m - Y/S/M
# #annual
# #do
# clim_doY_mmol_L_cal <- crop(clim_doY_mmol_L, e_cal)
# 
# #temp
# clim_temp_Y0_cal <- crop(clim_temp_Y0, e_cal_cl)
# 
# #seasonal
# #do
# clim_do_Wn0_cal <- crop(clim_do_Wn0, e_cal)
# clim_do_Sp0_cal <- crop(clim_do_Sp0, e_cal)
# clim_do_Su0_cal <- crop(clim_do_Su0, e_cal)
# clim_do_Fa0_cal <- crop(clim_do_Fa0, e_cal)
# 
# #temp
# clim_temp_Wn0_cal <- crop(clim_temp_Wn0, e_cal_cl)
# clim_temp_Sp0_cal <- crop(clim_temp_Sp0, e_cal_cl)
# clim_temp_Su0_cal <- crop(clim_temp_Su0, e_cal_cl)
# clim_temp_Fa0_cal <- crop(clim_temp_Fa0, e_cal_cl)
# 
# #month
# #do
# clim_doM_mmol_L_cal <- crop(clim_doM_mmol_L, e_cal)
# 
# #temp
# clim_temp_M0_cal <- crop(clim_temp_M0, e_cal_cl)
# 
# #Clim 200m - Y/S/M -- DO and Temp PENDING ON RESULTS
# #annual
# #seasonal
# #month
```

## DO (mmol/L) at 0m

### Mean across study period (2003-2015)

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_0m
#| fig-width: 10
#| fig-height: 8

#Mercator plot -- mean across all years
merc_do_Y0 <- aggregate(merc_do_Y0, fact = 4)
merc_do_Y0 <- mean(merc_do_Y0, na.rm=TRUE)
do1 <- ggplot() + 
  geom_spatraster(data = merc_do_Y0) + 
  scale_fill_viridis_c()+
    ggtitle("Mercator") +
  labs(fill = "DO (mmol/L)")+
  theme_minimal()

#WOA plot -- mean across all years
clim_do_Y0 <- mean(clim_doY_mmol_L, na.rm = TRUE)
do2 <- ggplot() + 
  geom_spatraster(data = clim_do_Y0) + 
  scale_fill_viridis_c()+
   labs(fill = "DO (mmol/L)")+
    ggtitle("WOA") +
  theme_minimal()

#calcofi plot -- mean across all years for each station 
calcofi_coord0 <- st_as_sf(calcofi_samp_coords, coords = c("Lon_Deg", "Lat_Deg"), crs = st_crs(4326))

calcofi_sta0 <- calcofi_coord0 %>% 
  group_by(geometry) %>% 
  summarise(mean_do0_mmol_L = mean(do0_mmol_L, na.rm = TRUE))

world <- ne_coastline(scale = "medium", returnclass = "sf")
st_crs(world) <- st_crs(4326)

do3 <- ggplot()+
  geom_sf(data = calcofi_sta0, aes(color = mean_do0_mmol_L), size = 3)+
  geom_sf(data = world)+
  coord_sf(xlim = c(-125, -117), ylim = c(29, 37)) +
  scale_color_viridis_c()+
   labs(fill = "DO (mmol/L)")+
    ggtitle("CalCOFI") +
  theme_minimal()

do1|do2|do3
```

### Mercator difference plots

```{r}
#| warning: false
#| message: false
#| label: DO_diff_0m
#| fig-width: 10
#| fig-height: 8
#| echo: false

#Merc vs. WOA
e <- ext(-153, -104, 2, 49)

#Merc vs. WOA
origin(merc_do_Y0) <- c(0,0)
merc_do_Y0 <- crop(merc_do_Y0, e)
clim_do_Y0 <- crop(clim_do_Y0, e)

diff_merc_clim <- diff(c(merc_do_Y0, clim_do_Y0))

d_do1 <- ggplot() + 
  geom_spatraster(data = diff_merc_clim) + 
  ggtitle("Comparing Mercator to WOA") +
  labs(fill = "Difference (mmol/L)") + 
  scale_fill_viridis_c(limits = c(-0.02, 0.04))+
  theme_minimal()

#Merc vs. CalCOFI
vect_cal_do0 <- vect(calcofi_sta0)
df_cal_do0 <- as.data.frame(vect_cal_do0)
merc_cal_extract <- extract(merc_do_Y0, vect_cal_do0)
diff_merc_cal <- merc_cal_extract$mean - df_cal_do0$mean_do0_mmol_L

calcofi_sta0$merc_diff <- diff_merc_cal

d_do2 <- ggplot()+
  geom_sf(data = calcofi_sta0, aes(color = merc_diff), size = 3)+
  geom_sf(data = world)+
  coord_sf(xlim = c(-125, -117), ylim = c(29, 37)) +
  scale_color_viridis_c(limits = c(-0.02, 0.04))+
  labs(color = "Difference (mmol/L)") + 
  ggtitle("Comparing Mercator to CalCOFI") +
  theme_minimal()

d_do1|d_do2
```

## DO (mmol/L) at 200m

### Mean across study period (2003-2015)

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_200m
#| fig-width: 10
#| fig-height: 6

#Mercator plot -- mean across all years
merc_do_Y200 <- aggregate(merc_do_Y200, fact = 4)
merc_do_Y200 <- mean(merc_do_Y200, na.rm=TRUE)

do1_200 <- ggplot() + 
  geom_spatraster(data = merc_do_Y200) + 
  scale_fill_viridis_c()+
  theme_minimal()+
    ggtitle("Mercator") +
  labs(fill = "DO (mmol/L)")

#WOA plot -- mean across all years
clim_do_Y200 <- mean(clim_doY_mmol_L200, na.rm = TRUE)
do2_200 <- ggplot() + 
  geom_spatraster(data = clim_do_Y200) + 
  scale_fill_viridis_c()+
   labs(fill = "DO (mmol/L)")+
    ggtitle("WOA") +
  theme_minimal()

#calcofi plot -- mean across all years for each station 
calcofi_coord200 <- st_as_sf(calcofi_samp_coords200, coords = c("Lon_Deg", "Lat_Deg"), crs = st_crs(4326))

calcofi_sta200 <- calcofi_coord200 %>% 
  group_by(geometry) %>% 
  summarise(mean_do200_mmol_L = mean(do200_mmol_L, na.rm = TRUE))

do3_200 <- ggplot()+
  geom_sf(data = calcofi_sta200, aes(color = mean_do200_mmol_L), size = 3)+
  geom_sf(data = world)+
  coord_sf(xlim = c(-125, -117), ylim = c(29, 37)) +
  scale_color_viridis_c()+
  theme_minimal()+
    ggtitle("CalCOFI") +
  labs(color = "DO (mmol/L)")

do1_200|do2_200|do3_200
```

### Mercator difference plots

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: DO_diff_200m
#| fig-width: 8
#| fig-height: 6
#Merc vs. WOA
origin(merc_do_Y200) <- c(0,0)
merc_do_Y200 <- crop(merc_do_Y200, e)
clim_do_Y200 <- crop(clim_do_Y200, e)

diff_merc_clim200 <- diff(c(merc_do_Y200, clim_do_Y200))

d_do1_200 <- ggplot() + 
  geom_spatraster(data = diff_merc_clim200) + 
  ggtitle("Comparing Mercator to WOA") +
  labs(fill = "Difference (mmol/L)") + 
  scale_fill_viridis_c(limits = c(-0.03, 0.10))+
  theme_minimal()

#Merc vs. CalCOFI
vect_cal_do200 <- vect(calcofi_sta200)
df_cal_do200 <- as.data.frame(vect_cal_do200)
merc_cal_extract200 <- extract(merc_do_Y200, vect_cal_do200)
diff_merc_cal200 <- merc_cal_extract200$mean - df_cal_do200$mean_do200_mmol_L

calcofi_sta200$merc_diff200 <- diff_merc_cal200

d_do2_200 <- ggplot()+
  geom_sf(data = calcofi_sta200, aes(color = merc_diff200), size = 3)+
  geom_sf(data = world)+
  coord_sf(xlim = c(-125, -117), ylim = c(29, 37)) +
  scale_color_viridis_c(limits = c(-0.03, 0.10))+ 
  labs(color = "Difference (mmol/L)") + 
  ggtitle("Comparing Mercator to CalCOFI") +
  theme_minimal()

d_do1_200|d_do2_200
```

## Temp (C) at 0m

::: panel-tabset
### Annual

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_0m_y

# merc_Y0 <- as.data.frame(global(merc_temp_Y0, "mean", na.rm=TRUE))
# colnames(merc_Y0) <- "Mercator means (C)"
# merc_Y0_cal <- as.data.frame(global(merc_temp_Y0_cal, "mean", na.rm=TRUE))
# 
# calcofi <- calcofi_samp0 %>% 
#   group_by(Year) %>% 
#   summarise(mean_temp0_C = mean(T_degC, na.rm = TRUE)) 
# 
# clim_Y0 <- as.data.frame(global(clim_temp_Y0, "mean", na.rm=TRUE))
# clim_Y0_cal <- as.data.frame(global(clim_temp_Y0_cal, "mean", na.rm=TRUE))
# 
# Temp_0_Y <- merc_Y0 %>%
#   mutate(`Merc calcofi domain (C)` = merc_Y0_cal$mean,
#          `OISST mean (C)` =  clim_Y0$mean, 
#          `OISST calcofi domain (C)` = clim_Y0_cal$mean,
#          `CalCOFI mean (C)` = calcofi$mean_temp0_C, 
#          `Merc vs. OISST` = `Mercator means (C)` - `OISST mean (C)`, 
#          `Merc vs. CalCOFI` = `Merc calcofi domain (C)` - `CalCOFI mean (C)`, 
#          `OISST vs. CalCOFI` = `OISST calcofi domain (C)`- `CalCOFI mean (C)`)
# 
# Temp_0_Y %>%
#   kbl() %>%
#   kable_minimal() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
```

### Seasonal

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_0m_s

# merc_0_s <- c(as.numeric(global(merc_temp_Wn0, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Sp0, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Su0, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Fa0, "mean", na.rm=TRUE)))
# 
# merc_0_s_cal <- c(as.numeric(global(merc_temp_Wn0_cal, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Sp0_cal, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Su0_cal, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Fa0_cal, "mean", na.rm=TRUE)))
# 
# clim_0_s <- c(as.numeric(global(clim_temp_Wn0, "mean", na.rm=TRUE)), as.numeric(global(clim_temp_Sp0, "mean", na.rm=TRUE)), as.numeric(global(clim_temp_Su0, "mean", na.rm=TRUE)), as.numeric(global(clim_temp_Fa0, "mean", na.rm=TRUE)))
# 
# clim_0_s_cal <- c(as.numeric(global(clim_temp_Wn0_cal, "mean", na.rm=TRUE)), as.numeric(global(clim_temp_Sp0_cal, "mean", na.rm=TRUE)), as.numeric(global(clim_temp_Su0_cal, "mean", na.rm=TRUE)), as.numeric(global(clim_temp_Fa0_cal, "mean", na.rm=TRUE)))
# 
# calcofi <- calcofi_samp0 %>% 
#   group_by(season) %>% 
#   summarise(mean_temp0_C = mean(T_degC, na.rm = TRUE)) %>%
#   mutate(season = as.factor(season))
# 
# levels(calcofi$season) <- c("Winter", "Spring", "Summer", "Fall")
# 
# Temp_0_S <- data.frame(Season = c("Winter", "Spring", "Summer", "Fall")) %>%
#   mutate(`Mercator means (C)` = merc_0_s, 
#          `Merc calcofi domain (C)` = merc_0_s_cal,
#          `OISST mean (C)` = clim_0_s, 
#          `OISST calcofi domain (C)` = clim_0_s_cal,
#          `CalCOFI mean (C)` = calcofi$mean_temp0_C, 
#          `Merc vs. OISST` = `Mercator means (C)` - `OISST mean (C)`, 
#          `Merc vs. CalCOFI` = `Merc calcofi domain (C)` - `CalCOFI mean (C)`, 
#          `OISST vs. CalCOFI` = `OISST calcofi domain (C)`- `CalCOFI mean (C)`)
# 
# Temp_0_S %>%
#   kbl() %>%
#   kable_minimal() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
```

### Monthly

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_0m_m
# merc_M0 <- as.data.frame(global(merc_temp_M0, "mean", na.rm=TRUE))
# colnames(merc_M0) <- "Mercator means (C)"
# merc_M0_cal <- as.data.frame(global(merc_temp_M0_cal, "mean", na.rm=TRUE))
# 
# calcofi <- calcofi_samp0 %>% 
#   group_by(Month) %>% 
#   summarise(mean_temp0_C = mean(T_degC, na.rm = TRUE)) %>%
#   add_row(Month = 12, mean_temp0_C = NA)
# 
# clim_M0 <- as.data.frame(global(clim_temp_M0, "mean", na.rm=TRUE))
# clim_M0_cal <- as.data.frame(global(clim_temp_M0_cal, "mean", na.rm=TRUE))
# 
# Temp_0_M <- merc_M0 %>%
#   mutate(`Merc calcofi domain (C)` = merc_M0_cal$mean,
#          `OISST mean (C)` =  clim_M0$mean, 
#          `OISST calcofi domain (C)` = clim_M0_cal$mean,
#          `CalCOFI mean (C)` = calcofi$mean_temp0_C, 
#          `Merc vs. OISST` = `Mercator means (C)` - `OISST mean (C)`, 
#          `Merc vs. CalCOFI` = `Merc calcofi domain (C)` - `CalCOFI mean (C)`, 
#          `OISST vs. CalCOFI` =  `OISST calcofi domain (C)` - `CalCOFI mean (C)`)
# 
# Temp_0_M %>%
#   kbl() %>%
#   kable_minimal() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
```
:::

## Temp (C) at 200m

This comparison is only available between Mercator and CalCOFI as OISST is only available at 0m depth. Depending on results, I can redo this using WOA data at 0m and at 200m.

::: panel-tabset
### Annual

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_200m_y

# merc_Y200 <- as.data.frame(global(merc_temp_Y200, "mean", na.rm=TRUE))
# colnames(merc_Y200) <- "Mercator means (C)"
# 
# merc_Y200_cal <- as.data.frame(global(merc_temp_Y200_cal, "mean", na.rm=TRUE))
# 
# calcofi <- calcofi_samp200 %>% 
#   group_by(Year) %>% 
#   summarise(mean_temp200_C = mean(T_degC, na.rm = TRUE)) 
# 
# Temp_200_Y <- merc_Y200 %>%
#   mutate(`Merc calcofi domian (C)` = merc_Y200_cal$mean,
#          `CalCOFI mean (C)` = calcofi$mean_temp200_C, 
#          `Merc vs. CalCOFI` = `Merc calcofi domian (C)` - `CalCOFI mean (C)`)
# 
# Temp_200_Y %>%
#   kbl() %>%
#   kable_minimal() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
```

### Seasonal

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_200m_s

# merc_200_s <- c(as.numeric(global(merc_temp_Wn200, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Sp200, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Su200, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Fa200, "mean", na.rm=TRUE)))
# 
# merc_200_s_cal <- c(as.numeric(global(merc_temp_Wn200_cal, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Sp200_cal, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Su200_cal, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Fa200_cal, "mean", na.rm=TRUE)))
# 
# calcofi <- calcofi_samp200 %>% 
#   group_by(season) %>% 
#   summarise(mean_temp200_C = mean(T_degC, na.rm = TRUE)) %>%
#   mutate(season = as.factor(season))
# 
# levels(calcofi$season) <- c("Winter", "Spring", "Summer", "Fall")
# 
# Temp_200_S <- data.frame(Season = c("Winter", "Spring", "Summer", "Fall")) %>%
#   mutate(`Merc calcofi domain (C)` = merc_200_s_cal,
#          `Mercator means (C)` = merc_200_s,
#          `CalCOFI mean (C)` = calcofi$mean_temp200_C,
#          `Merc vs. CalCOFI` = `Merc calcofi domain (C)` - `CalCOFI mean (C)`)
# 
# Temp_200_S %>%
#   kbl() %>%
#   kable_minimal() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))

```

### Monthly

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_200m_m

# merc_M200 <- as.data.frame(global(merc_temp_M200, "mean", na.rm=TRUE))
# colnames(merc_M200) <- "Mercator means (C)"
# 
# merc_M200_cal <- as.data.frame(global(merc_temp_M200_cal, "mean", na.rm=TRUE))
# 
# calcofi <- calcofi_samp200 %>% 
#   group_by(Month) %>% 
#   summarise(mean_temp200_C = mean(T_degC, na.rm = TRUE)) %>%
#   add_row(Month = 12, mean_temp200_C = NA)
# 
# Temp_200_M <- merc_M200 %>%
#   mutate(`Merc calcofi domain (C)` = merc_M200_cal$mean,
#          `CalCOFI mean (C)` = calcofi$mean_temp200_C, 
#          `Merc vs. CalCOFI` = `Merc calcofi domain (C)` - `CalCOFI mean (C)`)
# 
# Temp_200_M %>%
#   kbl() %>%
#   kable_minimal() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))

```
:::

TO DOOO
