---
title: "Mercator dissolved oxygen and temperature validation"
author: "Emily Nazario"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
code-block-bg: true
---

On this document I compare the Mercator BIORYS4 and GLORYS4 dissolved oxygen (DO) and temperature (temp) data to values from climatology products (DO: World Ocean Atlas; temp: OISST) and bottle collections at stations as part of CalCOFI and IMECOCAL. DO and temp values will be compared to these products as averages across the study area (lat: 1 to 49 and lon: -153 to -104) at 0m and at 200m. These spatial averages will also be compared on annual, seasonal, and monthly scales.

I should also note that the climatology data represent decadal averages either as a single raster over the entire product study time period (WOA - dissolved oxygen) by a given temporal resolution or separated by year and month (OISST - temperature).

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: libraries

### libraries #### 
library(tidyverse)
library(terra)
library(sf)
library(here)
library(ncdf4)
library(respR)
library(patchwork)
library(kableExtra)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Mercator data
## data ####
merc_do0_raw <- rast(here("data/enviro/model_validate/Mercator/CMEMS_CHL_DO_0m_JAN2003_Dec2015_0.25deg_D.nc"))
merc_do200_raw <- rast(here("data/enviro/model_validate/Mercator/CMEMS_CHL_DO_200m_JAN2003_Dec2015_0.25deg_D.nc"))
merc_temp0_raw <- list.files(here("data/enviro/model_validate/Mercator"), pattern = "Temp_0m", full.names = TRUE)
merc_temp200_raw <- list.files(here("data/enviro/model_validate/Mercator"), pattern = "Temp_200m", full.names = TRUE)

clim_do_ann <- rast(here("data/enviro/model_validate/Climatology/woa_oxygen/woa18_all_o00_01.nc"))
clim_do_mo <- list.files(here("data/enviro/model_validate/Climatology/woa_oxygen/monthly"), full.names = TRUE)
clim_tempWA <- rast(here("data/enviro/model_validate/Climatology/sst.mon.mean.nc"))

calcofi_samp <- read.csv(here("data/enviro/model_validate/CalCofi_IMECOCAL/CAlCOFI/194903-202105_Bottle.csv"), check.names = FALSE)
calcofi_met <- read.csv(here("data/enviro/model_validate/CalCofi_IMECOCAL/CAlCOFI/194903-202105_Cast.csv"), check.names = FALSE)
#imecocal

### Mercator data ####
#### oxygen (input: mmol/m^3, now: mmol/L) #### 
##### 0m #####
merc_do0_raw

#filter to separate variables by the "=" and prep to be able to filter by variable in next step
merc_vars0 <- str_split(names(merc_do0_raw), "=") %>% map_chr(\(x) x[1]) %>% str_extract("[^_]+")

#identify location of oxygen layers
merc_pos0 <- which(merc_vars0 == "o2")

#subset to only keep oxygen layers
merc_do0 <- merc_do0_raw %>% subset(merc_pos0) 
merc_do0 #spatial res: 0.25*, temporal es: daily, crs: lon/lat WGS 84

#convert DO to mmol/L
#convert temporal resolution to monthly and annually. Convert DO values from mmol/m^3 to mmol/L.
merc_do_M0 <- tapp(merc_do0, "months", mean)
merc_do_M0 <- merc_do_M0/1000
merc_do_Y0 <- tapp(merc_do0, "years", mean)
merc_do_Y0 <- merc_do_Y0/1000

#convert monthly resolution to seasonal 
#winter
merc_do_Wn0 <- subset(merc_do_M0, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
merc_do_Wn0 <- app(merc_do_Wn0, mean)
#spring
merc_do_Sp0 <- subset(merc_do_M0, c("m_3", "m_4", "m_5")) #Mar/Apr/May
merc_do_Sp0 <- app(merc_do_Sp0, mean)
#summer
merc_do_Su0 <- subset(merc_do_M0, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
merc_do_Su0 <- app(merc_do_Su0, mean)
#fall
merc_do_Fa0 <- subset(merc_do_M0, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
merc_do_Fa0 <- app(merc_do_Fa0, mean)

##### 200m #####
#filter to separate variables by the "=" and prep to be able to filter by variable in next step
merc_vars200 <- str_split(names(merc_do200_raw), "=") %>% map_chr(\(x) x[1]) %>% str_extract("[^_]+")

#identify location of oxygen layers
merc_pos200 <- which(merc_vars200 == "o2")

#subset to only keep oxygen layers
merc_do200 <- merc_do200_raw %>% subset(merc_pos200) 
merc_do200 #spatial res: 0.25*, temporal es: daily, crs: lon/lat WGS 84

#convert DO to mmol/L
#convert temporal resolution to monthly and annually. Convert DO from mmol/m^3 to mmol/L.
merc_do_M200 <- tapp(merc_do200, "months", mean)
merc_do_M200 <- merc_do_M200/1000
merc_do_Y200 <- tapp(merc_do200, "years", mean)
merc_do_Y200 <- merc_do_Y200/1000

#convert monthly resolution to seasonal 
#winter
merc_do_Wn200 <- subset(merc_do_M200, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
merc_do_Wn200 <- app(merc_do_Wn200, mean)
#spring
merc_do_Sp200 <- subset(merc_do_M200, c("m_3", "m_4", "m_5")) #Mar/Apr/May
merc_do_Sp200 <- app(merc_do_Sp200, mean)
#summer
merc_do_Su200 <- subset(merc_do_M200, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
merc_do_Su200 <- app(merc_do_Su200, mean)
#fall
merc_do_Fa200 <- subset(merc_do_M200, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
merc_do_Fa200 <- app(merc_do_Fa200, mean)

#### temperature (C) #### 
##### 0m #####
#combine all temp rasts into two rasters separated by depth layer

merc_temp_all0 <- NULL
for(i in 1:length(merc_temp0_raw)){
  # temp_nc <- nc_open(merc_temp0_raw[[i]])
  # 
  # #get correct lat and lon info from ncdf4 package
  # mod_rast <- rast(
  #   rast_lat <- ncvar_get(temp_nc, "nav_lat"),
  #   rast_lon <- ncvar_get(temp_nc, "nav_lon")
  # )
  # 
  # temp_lat <- rast_lat[1,1:241]
  # temp_lon <- rast_lon[1,1:241]
  
  temp_rast <- rast(merc_temp0_raw[[i]])
  rast_source <- sources(temp_rast)
  rast_year <- str_extract_all(rast_source, "\\d{4}")[[1]][1]
  
  d1 <- seq(as.Date(paste0(rast_year,"-","01-01")), by = "month",length = 6)
  d2 <- seq(as.Date(paste0(rast_year,"-","07-01")), by = "month",length = 6)
  
  #add year-month info to raster files
  if(str_detect(rast_source, "JAN") == TRUE){
    time(temp_rast) <- d1
  } else {
    time(temp_rast) <- d2
  }
  
  merc_temp_all0 <- append(merc_temp_all0, temp_rast)
}

merc_temp_all0 #spatial res: 0.25*, temporal es: year-month, crs: lon/lat WGS 84

#convert temporal resolution to monthly and annually
merc_temp_M0 <- tapp(merc_temp_all0, "months", mean)
merc_temp_Y0 <- tapp(merc_temp_all0, "years", mean)

#convert monthly resolution to seasonal 
#winter
merc_temp_Wn0 <- subset(merc_temp_M0, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
merc_temp_Wn0 <- app(merc_temp_Wn0, mean)
#spring
merc_temp_Sp0 <- subset(merc_temp_M0, c("m_3", "m_4", "m_5")) #Mar/Apr/May
merc_temp_Sp0 <- app(merc_temp_Sp0, mean)
#summer
merc_temp_Su0 <- subset(merc_temp_M0, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
merc_temp_Su0 <- app(merc_temp_Su0, mean)
#fall
merc_temp_Fa0 <- subset(merc_temp_M0, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
merc_temp_Fa0 <- app(merc_temp_Fa0, mean)

##### 200m #####
#combine all temp rasts into two rasters separated by depth layer
merc_temp_all200 <- NULL
for(i in 1:length(merc_temp200_raw)){
  temp_rast <- rast(merc_temp200_raw[[i]])
  rast_source <- sources(temp_rast)
  rast_year <- str_extract_all(rast_source, "\\d{4}")[[1]][1]
  
  d1 <- seq(as.Date(paste0(rast_year,"-","01-01")), by = "month",length = 6)
  d2 <- seq(as.Date(paste0(rast_year,"-","07-01")), by = "month",length = 6)
  
  #add year-month info to raster files
  if(str_detect(rast_source, "JAN") == TRUE){
    time(temp_rast) <- d1
  } else {
    time(temp_rast) <- d2
  }
  
  merc_temp_all200 <- append(merc_temp_all200, temp_rast)
}

merc_temp_all200 #spatial res: 0.25*, temporal es: year-month, crs: lon/lat WGS 84

#convert temporal resolution to monthly and annually
merc_temp_M200 <- tapp(merc_temp_all200, "months", mean)
merc_temp_Y200 <- tapp(merc_temp_all200, "years", mean)

#convert monthly resolution to seasonal 
#winter
merc_temp_Wn200 <- subset(merc_temp_M200, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
merc_temp_Wn200 <- app(merc_temp_Wn200, mean)
#spring
merc_temp_Sp200 <- subset(merc_temp_M200, c("m_3", "m_4", "m_5")) #Mar/Apr/May
merc_temp_Sp200 <- app(merc_temp_Sp200, mean)
#summer
merc_temp_Su200 <- subset(merc_temp_M200, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
merc_temp_Su200 <- app(merc_temp_Su200, mean)
#fall
merc_temp_Fa200 <- subset(merc_temp_M200, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
merc_temp_Fa200 <- app(merc_temp_Fa200, mean)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Climatology data

### Climatology data ####
#### oxygen (input: umol/kg, now: mmol/L) ####
#define extent to study area
e <- ext(-153, -104, 1, 49)
##### 0m #####
#decadal values for o2 conversion at surface for annual average
temp_o20 <- rast(here("data/enviro/model_validate/Climatology/woa_oxygen/temp_for_conversions/woa18_decav_t00_01.nc"))
temp_o20 <- crop(temp_o20,e)
temp_o20 <- temp_o20$`t_an_depth=0`

sal_o20 <- rast(here("data/enviro/model_validate/Climatology/woa_oxygen/sal_for_conversions/woa18_decav_s00_01.nc"))
sal_o20 <- crop(sal_o20,e)
sal_o20 <- sal_o20$`s_an_depth=0`

#spatial res: 0.1*, temporal res: averaged annually, temporal range: 1960-2017
#annual
clim_do_ann0W <- clim_do_ann$`o_mn_depth=0` 

clim_do_ann0 <- crop(clim_do_ann0W, e)
#plot(clim_do_ann0)

#convert to mmol/L
clim_doY_mmol_L <- clim_do_ann0*1.025*0.001 #1.025 kg/L and 0.001mmol in umol
#plot(clim_doY_mmol_L)

#monthly and seasonal files
clim_do_mo0 = NULL
for(i in 1:length(clim_do_mo)){
  temp_rast <- rast(clim_do_mo[[i]])
  temp_rast_mean <- temp_rast$`o_mn_depth=0`
  temp_rast_mean <- crop(temp_rast_mean, e)
  
  rast_name <- paste0("m", "_", i)
  names(temp_rast_mean) <- rast_name
  
  clim_do_mo0 <- append(clim_do_mo0, temp_rast_mean)
}

#monthly
clim_doM_mmol_L <- clim_do_mo0*1.025*0.001

#convert monthly resolution to seasonal 
#winter
clim_do_Wn0 <- subset(clim_doM_mmol_L, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
clim_do_Wn0 <- app(clim_do_Wn0, mean)
#spring
clim_do_Sp0 <- subset(clim_doM_mmol_L, c("m_3", "m_4", "m_5")) #Mar/Apr/May
clim_do_Sp0 <- app(clim_do_Sp0, mean)
#summer
clim_do_Su0 <- subset(clim_doM_mmol_L, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
clim_do_Su0 <- app(clim_do_Su0, mean)
#fall
clim_do_Fa0 <- subset(clim_doM_mmol_L, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
clim_do_Fa0 <- app(clim_do_Fa0, mean)

##### 200m #####
#spatial res: 0.1*, temporal res: averaged annually, temporal range: 1960-2017
clim_do_ann200W <- clim_do_ann$`o_mn_depth=200` 
clim_do_ann200 <- crop(clim_do_ann200W, e)

clim_doY_mmol_L200 <- clim_do_ann200*1.025*0.001
#plot(clim_doY_mmol_L200)

#need to do monthly and seasonal files
clim_do_mo200 = NULL
for(i in 1:length(clim_do_mo)){
  temp_rast <- rast(clim_do_mo[[i]])
  temp_rast_mean <- temp_rast$`o_mn_depth=200`
  temp_rast_mean <- crop(temp_rast_mean, e)
  
  rast_name <- paste0("m", "_", i)
  names(temp_rast_mean) <- rast_name
  
  clim_do_mo200 <- append(clim_do_mo200, temp_rast_mean)
}

clim_doM_mmol_L200 <- clim_do_mo200*1.025*0.001

#convert monthly resolution to seasonal 
#winter
clim_do_Wn200 <- subset(clim_doM_mmol_L200, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
clim_do_Wn200 <- app(clim_do_Wn200, mean)
global(clim_do_Wn200, "mean", na.rm=TRUE)

#spring
clim_do_Sp200 <- subset(clim_doM_mmol_L200, c("m_3", "m_4", "m_5")) #Mar/Apr/May
clim_do_Sp200 <- app(clim_do_Sp200, mean)
global(clim_do_Sp200, "mean", na.rm=TRUE)

#summer
clim_do_Su200 <- subset(clim_doM_mmol_L200, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
clim_do_Su200 <- app(clim_do_Su200, mean)
global(clim_do_Su200, "mean", na.rm=TRUE)

#fall
clim_do_Fa200 <- subset(clim_doM_mmol_L200, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
clim_do_Fa200 <- app(clim_do_Fa200, mean)
global(clim_do_Fa200, "mean", na.rm=TRUE)

#### temperature (C) ####
##### 0m #####
#clim_tempWA #spatial res: 0.25*, temporal res: annual, temporal range: 1981-2024
#plot(clim_tempWA)

e2 <- ext(207, 256, 1, 49)
clim_tempA <- crop(clim_tempWA, e2)
#plot(clim_tempA$sst_1)

clim_tempA <- subset(clim_tempA, time(clim_tempA) > as.Date("2003-01-01"))
clim_tempA <- subset(clim_tempA, time(clim_tempA) < as.Date("2015-12-31"))
#clim_tempA

#convert temporal resolution to monthly and annually
clim_temp_M0 <- tapp(clim_tempA, "months", mean)
clim_temp_Y0 <- tapp(clim_tempA, "years", mean)

#convert monthly resolution to seasonal 
#winter
clim_temp_Wn0 <- subset(clim_temp_M0, c("m_12", "m_1", "m_2")) #Dec/Jan/Feb
clim_temp_Wn0 <- app(clim_temp_Wn0, mean)
#spring
clim_temp_Sp0 <- subset(clim_temp_M0, c("m_3", "m_4", "m_5")) #Mar/Apr/May
clim_temp_Sp0 <- app(clim_temp_Sp0, mean)
#summer
clim_temp_Su0 <- subset(clim_temp_M0, c("m_6", "m_7", "m_8")) #Jun/Jul/Aug
clim_temp_Su0 <- app(clim_temp_Su0, mean)
#fall
clim_temp_Fa0 <- subset(clim_temp_M0, c("m_9", "m_10", "m_11")) #Sep/Oct/Nov
clim_temp_Fa0 <- app(clim_temp_Fa0, mean)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Station data

### Station data ####
#### CalCOFI ####
#filter to correct dates
calcofi_met_filt <- calcofi_met %>%
  filter(Year >= 2003 & Year <= 2015)

calcofi_samp_filt <- calcofi_samp[calcofi_samp$Cst_Cnt %in% calcofi_met_filt$Cst_Cnt, ]

#subset metadata to just Cst_Cnt ID, year, and month to join to sample DF
calcofi_met_sub <- calcofi_met_filt %>% 
  subset(select = c(Cst_Cnt, Year, Month))

calcofi_samp_filt1 <- calcofi_samp_filt %>%
  full_join(calcofi_met_sub, by = "Cst_Cnt")

calcofi_samp_filt1 <- calcofi_samp_filt1 %>%
  mutate(season = ifelse(Month >= 3 & Month <= 5, "Spring", 
                         ifelse(Month >= 6 & Month <= 8, "Summer", 
                                ifelse(Month >= 9 & Month <= 11, "Fall", "Winter"))))

##### 0m #####
calcofi_samp0 <- calcofi_samp_filt1 %>% 
  filter(Depthm == 0) %>%
  filter(!is.na(O2ml_L))

#convert DO to mmol/L for each row either using the average temp/sal if data isn't available, or the in situ measurements at that station/bottle collection
mean_sal = mean(calcofi_samp0$Salnty, na.rm = TRUE)
mean_temp = mean(calcofi_samp0$T_degC, na.rm = TRUE)
calcofi_samp0$do0_mmol_L = NA

for(i in 1:nrow(calcofi_samp0)){
  if(is.na(calcofi_samp0$Salnty[i]) | is.na(calcofi_samp0$T_degC[i])){
    
   calcofi_samp0$do0_mmol_L[i] = convert_DO(calcofi_samp0$O2ml_L[i], from = "mL/L", to = "mmol/L", S = mean_sal, t = mean_temp, P = 1.013253)
   
  } else if(!is.na(calcofi_samp0$Salnty[i]) & !is.na(calcofi_samp0$T_degC[i])){
    
    calcofi_samp0$do0_mmol_L[i] = convert_DO(calcofi_samp0$O2ml_L[i], from = "mL/L", to = "mmol/L", S = calcofi_samp0$Salnty[i], t = calcofi_samp0$T_degC[i], P = 1.013253)
    
  }
  #print(calcofi_samp0$do0_mmol_L[i])
}

##### 200m #####
calcofi_samp200 <- calcofi_samp_filt1 %>%
  filter(Depthm == 200) %>%
  filter(!is.na(O2ml_L))

```

## DO (mmol/L) at 0m

::: panel-tabset
### Annual

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_0m_y

merc_YO <- as.data.frame(global(merc_do_Y0, "mean", na.rm=TRUE))
colnames(merc_YO) <- "Mercator means (mmol/L)"

calcofi <- calcofi_samp0 %>% 
  group_by(Year) %>% 
  summarise(mean_do0_mmol_L = mean(do0_mmol_L, na.rm = TRUE))

DO_0_Y <- merc_YO %>%
  mutate(`WOA mean (mmol/L)` = as.numeric(global(clim_doY_mmol_L, "mean", na.rm=TRUE)), 
         `CalCOFI mean (mmol/L)` = calcofi$mean_do0_mmol_L, 
         `Merc vs. WOA` = `Mercator means (mmol/L)` - `WOA mean (mmol/L)`, 
         `Merc vs. CalCOFI` = `Mercator means (mmol/L)` - `CalCOFI mean (mmol/L)`, 
         `WOA vs. CalCOFI` = `WOA mean (mmol/L)`- `CalCOFI mean (mmol/L)`)

DO_0_Y %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Seasonal

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_0m_s

merc_0_s <- c(as.numeric(global(merc_do_Wn0, "mean", na.rm=TRUE)), as.numeric(global(merc_do_Sp0, "mean", na.rm=TRUE)), as.numeric(global(merc_do_Su0, "mean", na.rm=TRUE)), as.numeric(global(merc_do_Fa0, "mean", na.rm=TRUE)))

clim_0_s <- c(as.numeric(global(clim_do_Wn0, "mean", na.rm=TRUE)), as.numeric(global(clim_do_Sp0, "mean", na.rm=TRUE)), as.numeric(global(clim_do_Su0, "mean", na.rm=TRUE)), as.numeric(global(clim_do_Fa0, "mean", na.rm=TRUE)))

calcofi <- calcofi_samp0 %>% 
  group_by(season) %>% 
  summarise(mean_do0_mmol_L = mean(do0_mmol_L, na.rm = TRUE)) %>%
  mutate(season = as.factor(season))

levels(calcofi$season) <- c("Winter", "Spring", "Summer", "Fall")

DO_0_S <- data.frame(Season = c("Winter", "Spring", "Summer", "Fall")) %>%
  mutate(`Mercator means (mmol/L)` = merc_0_s, 
         `WOA mean (mmol/L)` = clim_0_s, 
         `CalCOFI mean (mmol/L)` = calcofi$mean_do0_mmol_L, 
         `Merc vs. WOA` = `Mercator means (mmol/L)` - `WOA mean (mmol/L)`, 
         `Merc vs. CalCOFI` = `Mercator means (mmol/L)` - `CalCOFI mean (mmol/L)`, 
         `WOA vs. CalCOFI` = `WOA mean (mmol/L)`- `CalCOFI mean (mmol/L)`)

DO_0_S %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Monthly

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_0m_m

merc_MO <- as.data.frame(global(merc_do_M0, "mean", na.rm=TRUE))
colnames(merc_MO) <- "Mercator means (mmol/L)"

calcofi <- calcofi_samp0 %>% 
  group_by(Month) %>% 
  summarise(mean_do0_mmol_L = mean(do0_mmol_L, na.rm = TRUE)) %>%
  add_row(Month = 12, mean_do0_mmol_L = NA)

clim <- as.data.frame(global(clim_doM_mmol_L, "mean", na.rm=TRUE))

DO_0_M <- merc_MO %>%
  mutate(`WOA mean (mmol/L)` =  clim$mean, 
         `CalCOFI mean (mmol/L)` = calcofi$mean_do0_mmol_L, 
         `Merc vs. WOA` = `Mercator means (mmol/L)` - `WOA mean (mmol/L)`, 
         `Merc vs. CalCOFI` = `Mercator means (mmol/L)` - `CalCOFI mean (mmol/L)`, 
         `WOA vs. CalCOFI` = `WOA mean (mmol/L)`- `CalCOFI mean (mmol/L)`)

DO_0_M %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```
:::

## DO (mmol/L) at 200m

This comparison is only available between Mercator and WOA as I was unsure how to convert the CalCOFI DO units (mL/L) at depth (200m) to mmol/L so they were comparable to the other data sources.

::: panel-tabset
### Annual

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_200m_y

merc_Y200 <- as.data.frame(global(merc_do_Y200, "mean", na.rm=TRUE))
colnames(merc_Y200) <- "Mercator means (mmol/L)"

DO_200_Y <- merc_Y200 %>%
  mutate(`WOA mean (mmol/L)` = as.numeric(global(clim_doY_mmol_L200, "mean", na.rm=TRUE)), 
         `Merc vs. WOA` = `Mercator means (mmol/L)` - `WOA mean (mmol/L)`)

DO_200_Y %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Seasonal

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_200m_s

merc_200_s <- c(as.numeric(global(merc_do_Wn200, "mean", na.rm=TRUE)), as.numeric(global(merc_do_Sp200, "mean", na.rm=TRUE)), as.numeric(global(merc_do_Su200, "mean", na.rm=TRUE)), as.numeric(global(merc_do_Fa200, "mean", na.rm=TRUE)))

clim_200_s <- c(as.numeric(global(clim_do_Wn200, "mean", na.rm=TRUE)), as.numeric(global(clim_do_Sp200, "mean", na.rm=TRUE)), as.numeric(global(clim_do_Su200, "mean", na.rm=TRUE)), as.numeric(global(clim_do_Fa200, "mean", na.rm=TRUE)))

DO_200_S <- data.frame(Season = c("Winter", "Spring", "Summer", "Fall")) %>%
  mutate(`Mercator means (mmol/L)` = merc_200_s, 
         `WOA mean (mmol/L)` = clim_200_s, 
         `Merc vs. WOA` = `Mercator means (mmol/L)` - `WOA mean (mmol/L)`)

DO_200_S %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Monthly

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_200m_m

merc_M20O <- as.data.frame(global(merc_do_M200, "mean", na.rm=TRUE))
colnames(merc_M20O) <- "Mercator means (mmol/L)"

clim <- as.data.frame(global(clim_doM_mmol_L200, "mean", na.rm=TRUE))

DO_200_M <- merc_M20O %>%
  mutate(`WOA mean (mmol/L)` =  clim$mean,
         `Merc vs. WOA` = `Mercator means (mmol/L)` - `WOA mean (mmol/L)`)

DO_200_M %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
:::

## Temp (C) at 0m

::: panel-tabset
### Annual

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_0m_y

merc_Y0 <- as.data.frame(global(merc_temp_Y0, "mean", na.rm=TRUE))
colnames(merc_Y0) <- "Mercator means (C)"

calcofi <- calcofi_samp0 %>% 
  group_by(Year) %>% 
  summarise(mean_temp0_C = mean(T_degC, na.rm = TRUE)) 

clim_Y0 <- as.data.frame(global(clim_temp_Y0, "mean", na.rm=TRUE))

Temp_0_Y <- merc_Y0 %>%
  mutate(`OISST mean (C)` =  clim_Y0$mean, 
         `CalCOFI mean (C)` = calcofi$mean_temp0_C, 
         `Merc vs. OISST` = `Mercator means (C)` - `OISST mean (C)`, 
         `Merc vs. CalCOFI` = `Mercator means (C)` - `CalCOFI mean (C)`, 
         `OISST vs. CalCOFI` = `OISST mean (C)`- `CalCOFI mean (C)`)

Temp_0_Y %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Seasonal

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_0m_s

merc_0_s <- c(as.numeric(global(merc_temp_Wn0, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Sp0, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Su0, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Fa0, "mean", na.rm=TRUE)))

clim_0_s <- c(as.numeric(global(clim_temp_Wn0, "mean", na.rm=TRUE)), as.numeric(global(clim_temp_Sp0, "mean", na.rm=TRUE)), as.numeric(global(clim_temp_Su0, "mean", na.rm=TRUE)), as.numeric(global(clim_temp_Fa0, "mean", na.rm=TRUE)))

calcofi <- calcofi_samp0 %>% 
  group_by(season) %>% 
  summarise(mean_temp0_C = mean(T_degC, na.rm = TRUE)) %>%
  mutate(season = as.factor(season))

levels(calcofi$season) <- c("Winter", "Spring", "Summer", "Fall")

Temp_0_S <- data.frame(Season = c("Winter", "Spring", "Summer", "Fall")) %>%
  mutate(`Mercator means (C)` = merc_0_s, 
         `OISST mean (C)` = clim_0_s, 
         `CalCOFI mean (C)` = calcofi$mean_temp0_C, 
         `Merc vs. OISST` = `Mercator means (C)` - `OISST mean (C)`, 
         `Merc vs. CalCOFI` = `Mercator means (C)` - `CalCOFI mean (C)`, 
         `OISST vs. CalCOFI` = `OISST mean (C)`- `CalCOFI mean (C)`)

Temp_0_S %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Monthly

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_0m_m
merc_M0 <- as.data.frame(global(merc_temp_M0, "mean", na.rm=TRUE))
colnames(merc_M0) <- "Mercator means (C)"

calcofi <- calcofi_samp0 %>% 
  group_by(Month) %>% 
  summarise(mean_temp0_C = mean(T_degC, na.rm = TRUE)) %>%
  add_row(Month = 12, mean_temp0_C = NA)

clim_M0 <- as.data.frame(global(clim_temp_M0, "mean", na.rm=TRUE))

Temp_0_M <- merc_M0 %>%
  mutate(`OISST mean (C)` =  clim_M0$mean, 
         `CalCOFI mean (C)` = calcofi$mean_temp0_C, 
         `Merc vs. OISST` = `Mercator means (C)` - `OISST mean (C)`, 
         `Merc vs. CalCOFI` = `Mercator means (C)` - `CalCOFI mean (C)`, 
         `OISST vs. CalCOFI` = `OISST mean (C)`- `CalCOFI mean (C)`)

Temp_0_M %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
:::

## Temp (C) at 200m

This comparison is only available between Mercator and CalCOFI as OISST is only available at 0m depth. Depending on results, I can redo this using WOA data at 0m and at 200m.

::: panel-tabset
### Annual

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_200m_y

merc_Y200 <- as.data.frame(global(merc_temp_Y200, "mean", na.rm=TRUE))
colnames(merc_Y200) <- "Mercator means (C)"

calcofi <- calcofi_samp200 %>% 
  group_by(Year) %>% 
  summarise(mean_temp200_C = mean(T_degC, na.rm = TRUE)) 

Temp_200_Y <- merc_Y200 %>%
  mutate(`CalCOFI mean (C)` = calcofi$mean_temp200_C, 
         `Merc vs. CalCOFI` = `Mercator means (C)` - `CalCOFI mean (C)`)

Temp_200_Y %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Seasonal

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_200m_s

merc_200_s <- c(as.numeric(global(merc_temp_Wn200, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Sp200, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Su200, "mean", na.rm=TRUE)), as.numeric(global(merc_temp_Fa200, "mean", na.rm=TRUE)))

calcofi <- calcofi_samp200 %>% 
  group_by(season) %>% 
  summarise(mean_temp200_C = mean(T_degC, na.rm = TRUE)) %>%
  mutate(season = as.factor(season))

levels(calcofi$season) <- c("Winter", "Spring", "Summer", "Fall")

Temp_200_S <- data.frame(Season = c("Winter", "Spring", "Summer", "Fall")) %>%
  mutate(`Mercator means (C)` = merc_200_s,
         `CalCOFI mean (C)` = calcofi$mean_temp200_C,
         `Merc vs. CalCOFI` = `Mercator means (C)` - `CalCOFI mean (C)`)

Temp_200_S %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

### Monthly

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: Temp_200m_m

merc_M200 <- as.data.frame(global(merc_temp_M200, "mean", na.rm=TRUE))
colnames(merc_M200) <- "Mercator means (C)"

calcofi <- calcofi_samp200 %>% 
  group_by(Month) %>% 
  summarise(mean_temp200_C = mean(T_degC, na.rm = TRUE)) %>%
  add_row(Month = 12, mean_temp200_C = NA)

Temp_200_M <- merc_M200 %>%
  mutate(`CalCOFI mean (C)` = calcofi$mean_temp200_C, 
         `Merc vs. CalCOFI` = `Mercator means (C)` - `CalCOFI mean (C)`)

Temp_200_M %>%
  kbl() %>%
  kable_minimal() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```
:::

## Remaining questions

### *\*pending if results look good enough to continue pursuing Mercator data products.*

1.  Do you have recommended approaches/code for converting units of CalCOFI DO data at depth to account for changes in pressure, temperature, and salinity?

2.  Should I expand temperature climatology analysis to 200m depth and instead get temperature data from WOA rather than OISST?

3.  Should I follow-up with the CalCOFI data and run an additional offshore vs. coastal comparison at each depth layer and temporal resolution?

4.  IMECOCAL data access.
