---
title: "Mercator dissolved oxygen and temperature validation"
author: "Emily Nazario"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
code-block-bg: true
---

On this document I compare the mean Mercator BIORYS4 and GLORYS4 dissolved oxygen (DO) and temperature (temp) data to values from climatology products (DO: World Ocean Atlas and CSIRO; temp: OISST (at 0m), CSIRO, and WOA) and CalCOFI stations. DO and temp values will be compared to the climatology products as averages across the study area (lat: 1 to 49 and lon: -153 to -104) at 0m and at 200m depth which represent two of the depth layers I will use for the shark models.

Temporal variability was compared by extracting the mean dissolved oxygen and temperature values at the nearest grid cell over the CalCOFI domain for each day a CalCOFI sample was performed throughout the study period (Jun 26th, 2003- Mar. 22nd, 2015). Clusters of coastal and offshore stations were selected (~5-10 stations for each zone) and the values were averaged across stations for each zone and month. Thus, the temporal variability plots capture the temperature and dissolved oxygen values on a monthly temporal resolution throughout the study's temporal range (2003-2015). This was repeated at 0m and 200m, and coastal vs. offshore data were plotted on separate panels. 

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: libraries

### libraries #### 
library(tidyverse)
library(terra)
library(sf)
library(here)
library(ncdf4)
library(respR)
library(patchwork)
library(kableExtra)
library(tidyterra)
library(rnaturalearth)
library(rnaturalearthdata)
library(colorRamps)
library(stars)
library(stringr)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Mercator data
## data ####
merc_do0_raw <- rast(here("data/enviro/model_validate/Mercator/CMEMS_CHL_DO_0m_JAN2003_Dec2015_0.25deg_D.nc"))
merc_do200_raw <- rast(here("data/enviro/model_validate/Mercator/CMEMS_CHL_DO_200m_JAN2003_Dec2015_0.25deg_D.nc"))
merc_temp0_raw <- list.files(here("data/enviro/model_validate/Mercator"), pattern = "Temp_0m", full.names = TRUE)
merc_temp200_raw <- list.files(here("data/enviro/model_validate/Mercator"), pattern = "Temp_200m", full.names = TRUE)

clim_do_ann <- rast(here("data/enviro/model_validate/Climatology/woa_oxygen/woa18_all_o00_01.nc"))
clim_do_mo <- list.files(here("data/enviro/model_validate/Climatology/woa_oxygen/monthly"), full.names = TRUE)
clim_tempWA <- rast(here("data/enviro/model_validate/Climatology/sst.mon.mean.nc"))

calcofi_samp <- read.csv(here("data/enviro/model_validate/CalCofi_IMECOCAL/CAlCOFI/194903-202105_Bottle.csv"), check.names = FALSE)
calcofi_met <- read.csv(here("data/enviro/model_validate/CalCofi_IMECOCAL/CAlCOFI/194903-202105_Cast.csv"), check.names = FALSE)
#imecocal

### Mercator data ####
#### oxygen (input: mmol/m^3, now: mmol/L) #### 
##### 0m #####
merc_do0_raw

#filter to separate variables by the "=" and prep to be able to filter by variable in next step
merc_vars0 <- str_split(names(merc_do0_raw), "=") %>% map_chr(\(x) x[1]) %>% str_extract("[^_]+")

#identify location of oxygen layers
merc_pos0 <- which(merc_vars0 == "o2")

#subset to only keep oxygen layers
merc_do0 <- merc_do0_raw %>% subset(merc_pos0) 
merc_do0 #spatial res: 0.25*, temporal es: daily, crs: lon/lat WGS 84

#convert DO to mmol/L
#convert temporal resolution to monthly and annually. Convert DO values from mmol/m^3 to mmol/L.
merc_do_M0 <- tapp(merc_do0, "months", mean)
merc_do_M0 <- merc_do_M0/1000
merc_do_Y0 <- tapp(merc_do0, "years", mean)
merc_do_Y0 <- merc_do_Y0/1000

##### 200m #####
#filter to separate variables by the "=" and prep to be able to filter by variable in next step
merc_vars200 <- str_split(names(merc_do200_raw), "=") %>% map_chr(\(x) x[1]) %>% str_extract("[^_]+")

#identify location of oxygen layers
merc_pos200 <- which(merc_vars200 == "o2")

#subset to only keep oxygen layers
merc_do200 <- merc_do200_raw %>% subset(merc_pos200) 
merc_do200 #spatial res: 0.25*, temporal es: daily, crs: lon/lat WGS 84

#convert DO to mmol/L
#convert temporal resolution to monthly and annually. Convert DO from mmol/m^3 to mmol/L.
merc_do_M200 <- tapp(merc_do200, "months", mean)
merc_do_M200 <- merc_do_M200/1000
merc_do_Y200 <- tapp(merc_do200, "years", mean)
merc_do_Y200 <- merc_do_Y200/1000

#### temperature (C) #### 
##### 0m #####
#combine all temp rasts into two rasters separated by depth layer. Fix temporal and spatial data.
merc_temp_all0 <- readRDS(here("data/enviro/model_validate/Mercator/processed/merc_temp_all0.rds"))

#convert temporal resolution to monthly and annually
merc_temp_M0 <- tapp(merc_temp_all0, "months", mean)
merc_temp_Y0 <- tapp(merc_temp_all0, "years", mean)

##### 200m #####
#combine all temp rasts into two rasters separated by depth layer
merc_temp_all200 <- readRDS(here("data/enviro/model_validate/Mercator/processed/merc_temp_all200.rds"))

#convert temporal resolution to monthly and annually
merc_temp_M200 <- tapp(merc_temp_all200, "months", mean)
merc_temp_Y200 <- tapp(merc_temp_all200, "years", mean)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Climatology data

### Climatology data ####
#### oxygen (input: umol/kg, now: mmol/L) ####
#define extent to study area
e <- ext(-153, -104, 1, 49)
##### 0m #####
#spatial res: 0.1*, temporal res: averaged annually, temporal range: 1960-2017
#annual
clim_do_ann0W <- clim_do_ann$`o_mn_depth=0` 
clim_do_ann0 <- crop(clim_do_ann0W, e)

#convert to mmol/L
clim_doY_mmol_L <- clim_do_ann0*1.025*0.001 #1.025 kg/L and 0.001mmol in umol

#monthly
clim_do_mo0 = NULL
for(i in 1:length(clim_do_mo)){
  temp_rast <- rast(clim_do_mo[[i]])
  temp_rast_mean <- temp_rast$`o_mn_depth=0`
  temp_rast_mean <- crop(temp_rast_mean, e)
  
  rast_name <- paste0("m", "_", i)
  names(temp_rast_mean) <- rast_name
  
  clim_do_mo0 <- append(clim_do_mo0, temp_rast_mean)
}

#monthly
clim_doM_mmol_L <- clim_do_mo0*1.025*0.001

##### 200m #####
#spatial res: 0.1*, temporal res: averaged annually, temporal range: 1960-2017
clim_do_ann200W <- clim_do_ann$`o_mn_depth=200` 
clim_do_ann200 <- crop(clim_do_ann200W, e)

clim_doY_mmol_L200 <- clim_do_ann200*1.025*0.001
#plot(clim_doY_mmol_L200)

#need to do monthly and seasonal files
clim_do_mo200 = NULL
for(i in 1:length(clim_do_mo)){
  temp_rast <- rast(clim_do_mo[[i]])
  temp_rast_mean <- temp_rast$`o_mn_depth=200`
  temp_rast_mean <- crop(temp_rast_mean, e)
  
  rast_name <- paste0("m", "_", i)
  names(temp_rast_mean) <- rast_name
  
  clim_do_mo200 <- append(clim_do_mo200, temp_rast_mean)
}

clim_doM_mmol_L200 <- clim_do_mo200*1.025*0.001

#### temperature (C) ####
##### 0m #####
#clim_tempWA #spatial res: 0.25*, temporal res: annual, temporal range: 1981-2024
e2 <- ext(207, 256, 1, 49)
clim_tempA <- crop(clim_tempWA, e2)

clim_tempA <- subset(clim_tempA, time(clim_tempA) > as.Date("2003-01-01"))
clim_tempA <- subset(clim_tempA, time(clim_tempA) < as.Date("2015-12-31"))

#convert temporal resolution to monthly and annually
clim_temp_M0 <- tapp(clim_tempA, "months", mean)
clim_temp_Y0 <- tapp(clim_tempA, "years", mean)

#WOA temp (0m and 200m)
temp_woa <- rast(here("data/enviro/model_validate/Climatology/woa_temp/woa18_decav_t00_01.nc"))
temp_woa <- crop(temp_woa,e)
temp_woa0 <- temp_woa$`t_mn_depth=0`
temp_woa200 <- temp_woa$`t_mn_depth=200`
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Station data

### Station data ####
#### CalCOFI ####
#filter to correct dates
calcofi_met_filt <- calcofi_met %>%
  filter(Year >= 2003 & Year <= 2015)

calcofi_samp_filt <- calcofi_samp[calcofi_samp$Cst_Cnt %in% calcofi_met_filt$Cst_Cnt, ]

#subset metadata to just Cst_Cnt ID, year, and month to join to sample DF
calcofi_met_sub <- calcofi_met_filt %>% 
  subset(select = c(Cst_Cnt, Year, Month, Date))

calcofi_samp_filt1 <- calcofi_samp_filt %>%
  full_join(calcofi_met_sub, by = "Cst_Cnt")

calcofi_samp_filt1 <- calcofi_samp_filt1 %>%
  mutate(season = ifelse(Month >= 3 & Month <= 5, "Spring", 
                         ifelse(Month >= 6 & Month <= 8, "Summer", 
                                ifelse(Month >= 9 & Month <= 11, "Fall", "Winter"))))

##### 0m #####
calcofi_samp0 <- calcofi_samp_filt1 %>% 
  filter(Depthm == 0) %>%
  filter(!is.na(O2ml_L)) %>%
  mutate(do0_mmol_L = O2ml_L/22.3916)

##### 200m #####
calcofi_samp200 <- calcofi_samp_filt1 %>%
  filter(Depthm == 200) %>%
  filter(!is.na(O2ml_L))%>%
  mutate(do200_mmol_L = O2ml_L/22.3916)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: CSIRO data

### Dissolved oxygen ####
cars_do <- nc_open(here("data/enviro/model_validate/CSIRO/oxygen_cars2009.nc"))

do_cars <- ncvar_get(cars_do, varid = "mean")
mean_oxy_var <- "mean"

lat_do <- ncvar_get(cars_do, "lat")
lon_do <- ncvar_get(cars_do, "lon")
lon_do[lon_do > 180] <- lon_do[lon_do > 180] - 360 #change from 0 to 360 to -180 to 180
order_lon <- order(lon_do)
lon_do <- lon_do[order_lon]

depths_do <- ncvar_get(cars_do, varid = "depth")
mean_oxy_data <- ncvar_get(cars_do, mean_oxy_var)

### 0m 
depth_0 <- which(depths_do == 0)
mean_oxy_0 <- mean_oxy_data[,,depth_0] #extract mean oxygen data at 0m assuming the variable has dimensions lon, lat, depth
mean_oxy_0 <- mean_oxy_0[order_lon,]

oxy_rast_0 <- rast(nrows = length(lat_do), ncols = length(lon_do), 
                   xmin = -180, xmax = 180, 
                   ymin = min(lat_do), ymax = max(lat_do))

values(oxy_rast_0) <- t(mean_oxy_0)
crs(oxy_rast_0) <- "EPSG:4326"
oxy_rast_0 <- flip(oxy_rast_0, direction = "vertical")

    #crop to study domain
e <- ext(-153, -104, 2, 49)
cars_do_0 <- crop(oxy_rast_0, e)

resamp_geom_cars <- rast(ext(-153, -104, 2, 49), 
                    crs = "EPSG:4326",
                    res = 0.50)
cars_do_resamp0 <- terra::resample(cars_do_0, resamp_geom_cars, method='bilinear')

    #convert oxygen values from ml/L to mmol/L
cars_do_resamp0 <- cars_do_resamp0/22.3916

### 200m 
depth_200 <- which(depths_do == 200)
mean_oxy_200 <- mean_oxy_data[,,depth_200] #extract mean oxygen data at 200m assuming the variable has dimensions lon, lat, depth
mean_oxy_200 <- mean_oxy_200[order_lon,]

oxy_rast_200 <- rast(nrows = length(lat_do), ncols = length(lon_do), 
                   xmin = -180, xmax = 180, 
                   ymin = min(lat_do), ymax = max(lat_do))

values(oxy_rast_200) <- t(mean_oxy_200)
crs(oxy_rast_200) <- "EPSG:4326"
oxy_rast_200 <- flip(oxy_rast_200, direction = "vertical")

    #crop to study domain
cars_do_200 <- crop(oxy_rast_200, e)
cars_do_resamp200 <- terra::resample(cars_do_200, resamp_geom_cars, method='bilinear')

    #convert oxygen values from ml/L to mmol/L
cars_do_resamp200 <- cars_do_resamp200/22.3916


### Temperature ####
cars_temp <- nc_open(here("data/enviro/model_validate/CSIRO/temperature_cars2009a.nc"))

temp_cars <- ncvar_get(cars_temp, varid = "mean")
mean_temp_var <- "mean"

lat_temp <- ncvar_get(cars_temp, "lat")
lon_temp <- ncvar_get(cars_temp, "lon")
lon_temp[lon_temp > 180] <- lon_temp[lon_temp > 180] - 360 #change from 0 to 360 to -180 to 180
order_lon <- order(lon_temp)
lon_temp <- lon_temp[order_lon]

depths_temp <- ncvar_get(cars_temp, varid = "depth")
mean_temp_data <- ncvar_get(cars_temp, mean_temp_var)

### 0m 
depth_0 <- which(depths_temp == 0)
mean_temp_0 <- mean_temp_data[,,depth_0] #extract mean oxygen data at 0m assuming the variable has dimensions lon, lat, depth
mean_temp_0 <- mean_temp_0[order_lon,]

temp_rast_0 <- rast(nrows = length(lat_temp), ncols = length(lon_temp), 
                   xmin = -180, xmax = 180, 
                   ymin = min(lat_temp), ymax = max(lat_temp))

values(temp_rast_0) <- t(mean_temp_0)
crs(temp_rast_0) <- "EPSG:4326"
temp_rast_0 <- flip(temp_rast_0, direction = "vertical")

    #crop to study domain
cars_temp_0 <- crop(temp_rast_0, e)
cars_temp_resamp0 <- terra::resample(cars_temp_0, resamp_geom_cars, method='bilinear')

### 200m
depth_200 <- which(depths_temp == 200)
mean_temp_200 <- mean_temp_data[,,depth_200] #extract mean oxygen data at 0m assuming the variable has dimensions lon, lat, depth
mean_temp_200 <- mean_temp_200[order_lon,]

temp_rast_200 <- rast(nrows = length(lat_temp), ncols = length(lon_temp), 
                   xmin = -180, xmax = 180, 
                   ymin = min(lat_temp), ymax = max(lat_temp))

values(temp_rast_200) <- t(mean_temp_200)
crs(temp_rast_200) <- "EPSG:4326"
temp_rast_200 <- flip(temp_rast_200, direction = "vertical")

    #crop to study domain
cars_temp_200 <- crop(temp_rast_200, e)
cars_temp_resamp200 <- terra::resample(cars_temp_200, resamp_geom_cars, method='bilinear')

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
#| label: Modifying merc and clim domains to calcofi

#subset metadata to just Cst_Cnt ID, year, and month to join to sample DF
calcofi_coords <- calcofi_met_filt %>% 
  subset(select = c(Cst_Cnt, Lat_Deg, Lon_Deg))

calcofi_samp_coords <- calcofi_samp0 %>%
  full_join(calcofi_coords, by = "Cst_Cnt")

calcofi_samp_coords200 <- calcofi_samp200 %>%
  full_join(calcofi_coords, by = "Cst_Cnt")
# min(calcofi_samp_coords$Lat_Deg)
# max(calcofi_samp_coords$Lat_Deg)
# min(calcofi_samp_coords$Lon_Deg)
# max(calcofi_samp_coords$Lon_Deg)

e_cal <- ext(-125, -117, 29, 37)
e_cal_m <- rast(xmin=650, xmax=680, ymin=621, ymax=660, nrows=25, ncols=25)
e_cal_cl <- ext(235, 243, 29, 37)

#Merc 0m - Y/S/M
#annual 
#do
#merc_do_Y0_cal <- crop(merc_do_Y0, e_cal)

#temp
#merc_temp_Y0_cal <- crop(merc_temp_Y0, e_cal_m)

# #month
# #do
# merc_do_M0_cal <- crop(merc_do_M0, e_cal)
# 
# #temp
# merc_temp_M0_cal <- crop(merc_temp_M0, e_cal_m)
# 
# #Merc 200m - Y/S/M -- DO PENDING ON RESULTS
# #annual
# #temp 
# merc_temp_Y200_cal <- crop(merc_temp_Y200, e_cal_m)
# #month
# #temp
# merc_temp_M200_cal <- crop(merc_temp_M200, e_cal_m)
# 
# 
# #Clim 0m - Y/S/M
# #annual
# #do
# clim_doY_mmol_L_cal <- crop(clim_doY_mmol_L, e_cal)
# 
# #temp
# clim_temp_Y0_cal <- crop(clim_temp_Y0, e_cal_cl)
# #temp
# clim_temp_Wn0_cal <- crop(clim_temp_Wn0, e_cal_cl)
# clim_temp_Sp0_cal <- crop(clim_temp_Sp0, e_cal_cl)
# clim_temp_Su0_cal <- crop(clim_temp_Su0, e_cal_cl)
# clim_temp_Fa0_cal <- crop(clim_temp_Fa0, e_cal_cl)
# 
# #month
# #do
# clim_doM_mmol_L_cal <- crop(clim_doM_mmol_L, e_cal)
# 
# #temp
# clim_temp_M0_cal <- crop(clim_temp_M0, e_cal_cl)
# 
# #Clim 200m - Y/S/M -- DO and Temp PENDING ON RESULTS
# #annual
# #seasonal
# #month
```

## DO (mmol/L) at 0m

### Mean across study period (2003-2015)

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_0m
#| fig-width: 10
#| fig-height: 8
#| fig-align: center

#Mercator plot -- mean across all years
merc_do_Y0_woa <- aggregate(merc_do_Y0, fact = 4)
merc_do_Y0_cars <- aggregate(merc_do_Y0, fact = 2)
merc_do_Y0_cars <- mean(merc_do_Y0_cars, na.rm = TRUE)
merc_do_Y0_woa <- mean(merc_do_Y0_woa, na.rm=TRUE)
merc_do_Y0_cal <- mean(merc_do_Y0, na.rm = TRUE)

do1 <- ggplot() + 
  geom_spatraster(data = merc_do_Y0_cal) + 
  scale_fill_gradientn(colours = matlab.like(30))+
    ggtitle("Mercator") +
  labs(fill = "DO (mmol/L)")+
  theme_minimal()

#WOA plot -- mean across all years
clim_do_Y0 <- mean(clim_doY_mmol_L, na.rm = TRUE)
do2 <- ggplot() + 
  geom_spatraster(data = clim_do_Y0) + 
  scale_fill_gradientn(colours = matlab.like(30))+
   labs(fill = "DO (mmol/L)")+
    ggtitle("WOA") +
  theme_minimal()

#CARS plot -- mean across all years (climtology)
do2.5 <- ggplot() + 
  geom_spatraster(data = cars_do_resamp0) + 
  scale_fill_gradientn(colours = matlab.like(30))+
   labs(fill = "DO (mmol/L)")+
    ggtitle("CSIRO") +
  theme_minimal()

#calcofi plot -- mean across all years for each station 
calcofi_coord0 <- st_as_sf(calcofi_samp_coords, coords = c("Lon_Deg", "Lat_Deg"), crs = st_crs(4326))

calcofi_sta0 <- calcofi_coord0 %>% 
  group_by(geometry) %>% 
  summarise(mean_do0_mmol_L = mean(do0_mmol_L, na.rm = TRUE))

world <- ne_coastline(scale = "medium", returnclass = "sf")
st_crs(world) <- st_crs(4326)

do3 <- ggplot()+
  geom_sf(data = calcofi_sta0, aes(color = mean_do0_mmol_L), size = 3)+
  scale_colour_gradientn(colours = matlab.like(30))+
   labs(color = "DO (mmol/L)")+
    ggtitle("CalCOFI") +
  theme_minimal()

ggpubr::ggarrange(do1, do2, do2.5, do3)
```

### Mercator difference plots

```{r}
#| warning: false
#| message: false
#| label: DO_diff_0m
#| fig-width: 10
#| fig-height: 8
#| echo: false
#| fig-align: center

#Merc vs. WOA
e <- ext(-153, -104, 2, 49)

origin(merc_do_Y0_woa) <- c(0,0)
merc_do_Y0_woa <- crop(merc_do_Y0_woa, e)
clim_do_Y0 <- crop(clim_do_Y0, e)

diff_merc_clim <- diff(c(merc_do_Y0_woa, clim_do_Y0))

d_do1 <- ggplot() + 
  geom_spatraster(data = diff_merc_clim) + 
  ggtitle("Comparing Mercator to WOA") +
  labs(fill = "Difference (mmol/L)") + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()

#Merc vs. CARS CSIRO
origin(merc_do_Y0_cars) <- c(0,0)
merc_do_Y0_cars <- crop(merc_do_Y0_cars, e)

diff_merc_cars <- diff(c(merc_do_Y0_cars, cars_do_resamp0))

d_do1.5 <- ggplot() + 
  geom_spatraster(data = diff_merc_cars) + 
  ggtitle("Comparing Mercator to CSIRO") +
  labs(fill = "Difference (mmol/L)") + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()

#Merc vs. CalCOFI
vect_cal_do0 <- vect(calcofi_sta0)
df_cal_do0 <- as.data.frame(vect_cal_do0)
merc_cal_extract <- terra::extract(merc_do_Y0_cal, vect_cal_do0)
diff_merc_cal <- merc_cal_extract$mean - df_cal_do0$mean_do0_mmol_L

calcofi_sta0$merc_diff <- diff_merc_cal

d_do2 <- ggplot()+
  geom_sf(data = calcofi_sta0, aes(color = merc_diff), size = 3)+
  scale_colour_gradientn(colours = matlab.like(30))+
  labs(color = "Difference (mmol/L)") + 
  ggtitle("Comparing Mercator to CalCOFI") +
  theme_minimal()

ggpubr::ggarrange(d_do1, d_do1.5, d_do2)
```

### Mercator temporal variation vs. CalCOFI data 0m

Here, the mercator dissolved oxygen rasters were subset the month-day-year that a CalCOFI bottle sample was collected. The plotted mercator values represent the closest grid cell to that CalCOFI station when a sample was collected. A daily temporal resolution was possible here.

```{r}
#| warning: false
#| message: false
#| echo: false
#| fig-align: center
#| fig-width: 12
#| fig-height: 6
#| label: temporal_variation0_DO

#Merc subset to CalCOFI domain in year-month temporal resolution 
merc_do_D0 <- merc_do0/1000
#merc_do_cal0 <- crop(merc_do_D0, e_cal)
calcofi_coord0$Date <- as.Date(calcofi_coord0$Date, format = "%m/%d/%Y")
calcofi_coord0 <- calcofi_coord0 %>% drop_na(Date)

merc_vari0 <- NULL
for(i in 1:nrow(calcofi_coord0)){
  #subset raster according to calcofi sample date
  temp_date <- calcofi_coord0$Date[i]
  temp_rast <- merc_do_D0[[time(merc_do_D0) == temp_date]]
  
  #get coords of station 
  sta_loc <- calcofi_coord0$geometry[i]
  
  #extract mercator temp at the cell closest to that station
  sta_vect <- vect(sta_loc)
  merc_extract <- terra::extract(temp_rast, sta_vect)
  
  #save merc cell value, station value, and difference between in three cols
  temp_vari <- NULL
  temp_vari$Date <- temp_date
  temp_vari$geometry <- sta_loc
  temp_vari$merc_do_0m <- merc_extract[,2]
  temp_vari$calcofi_do_0m <- calcofi_coord0$do0_mmol_L[i]
  
  df_vari <- as.data.frame(temp_vari)
  
  merc_vari0 <- rbind(merc_vari0, df_vari)
}

merc_vari0_l <- merc_vari0 %>% pivot_longer(c(3:4), names_to = "data_source", values_to = "DO_mmol_L") %>%
  mutate(sta_geom = as.character(geometry)) %>% st_as_sf()

#filter for coastal cluster of stations 
merc_vari0_coast <- st_crop(merc_vari0_l, xmin = -121, xmax = -117, ymax = 34, ymin = 31)
merc_vari0_coast <- merc_vari0_coast %>% 
  mutate(yr_mo = substr(Date, 1, 7)) %>%
  group_by(yr_mo, data_source) %>% 
  mutate(mean_DO = mean(DO_mmol_L, na.rm = TRUE))

ggplot(merc_vari0_coast, aes(Date, mean_DO, color = data_source)) + 
  geom_point(size = 2) + 
  geom_line() +
  scale_color_manual(values = c("steelblue4", "goldenrod"))+
  labs(subtitle = "Coastal stations") + 
  theme_minimal()

#filter for offshore cluster of stations 
merc_vari0_off <- st_crop(merc_vari0_l, xmin = -125, xmax = -121.5, ymax = 32.5, ymin = 28)
merc_vari0_off <- merc_vari0_off %>% 
  mutate(yr_mo = substr(Date, 1, 7)) %>%
  group_by(yr_mo, data_source) %>% 
  mutate(mean_DO = mean(DO_mmol_L, na.rm = TRUE))

ggplot(merc_vari0_off, aes(Date, mean_DO, color = data_source)) + 
  geom_point(size = 2) + 
  geom_line() +
  scale_color_manual(values = c("steelblue4", "goldenrod"))+
  labs(subtitle = "Offshore stations") + 
  theme_minimal()

```

## DO (mmol/L) at 200m

### Mean across study period (2003-2015)

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: DO_200m
#| fig-width: 10
#| fig-height: 8
#| fig-align: center

#Mercator plot -- mean across all years
merc_do_Y200_woa <- aggregate(merc_do_Y200, fact = 4)
merc_do_Y200_woa <- mean(merc_do_Y200_woa, na.rm=TRUE)
merc_do_Y200_cal <- mean(merc_do_Y200, na.rm = TRUE)
merc_do_Y200_cars <- aggregate(merc_do_Y200, fact = 2)
merc_do_Y200_cars <- mean(merc_do_Y200_cars, na.rm = TRUE)

do1_200 <- ggplot() + 
  geom_spatraster(data = merc_do_Y200_cal) + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()+
    ggtitle("Mercator") +
  labs(fill = "DO (mmol/L)")

#WOA plot -- mean across all years
clim_do_Y200 <- mean(clim_doY_mmol_L200, na.rm = TRUE)
do2_200 <- ggplot() + 
  geom_spatraster(data = clim_do_Y200) + 
  scale_fill_gradientn(colours = matlab.like(30))+
   labs(fill = "DO (mmol/L)")+
    ggtitle("WOA") +
  theme_minimal()

#CARS plot -- mean across all years (climtology)
do2.5_200 <- ggplot() + 
  geom_spatraster(data = cars_do_resamp200) + 
  scale_fill_gradientn(colours = matlab.like(30))+
   labs(fill = "DO (mmol/L)")+
    ggtitle("CSIRO") +
  theme_minimal()

#calcofi plot -- mean across all years for each station 
calcofi_coord200 <- st_as_sf(calcofi_samp_coords200, coords = c("Lon_Deg", "Lat_Deg"), crs = st_crs(4326))

calcofi_sta200 <- calcofi_coord200 %>% 
  group_by(geometry) %>% 
  summarise(mean_do200_mmol_L = mean(do200_mmol_L, na.rm = TRUE))

do3_200 <- ggplot()+
  geom_sf(data = calcofi_sta200, aes(color = mean_do200_mmol_L), size = 3)+
  scale_colour_gradientn(colours = matlab.like(30))+
  theme_minimal()+
    ggtitle("CalCOFI") +
  labs(color = "DO (mmol/L)")

ggpubr::ggarrange(do1_200,do2_200,do2.5_200,do3_200)
```

### Mercator difference plots

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: DO_diff_200m
#| fig-width: 10
#| fig-height: 8
#Merc vs. WOA
origin(merc_do_Y200_woa) <- c(0,0)
merc_do_Y200_woa <- crop(merc_do_Y200_woa, e)
clim_do_Y200 <- crop(clim_do_Y200, e)

diff_merc_clim200 <- diff(c(merc_do_Y200_woa, clim_do_Y200))

d_do1_200 <- ggplot() + 
  geom_spatraster(data = diff_merc_clim200) + 
  ggtitle("Comparing Mercator to WOA") +
  labs(fill = "Difference (mmol/L)") + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()

#Merc vs. CARS CSIRO
origin(merc_do_Y200_cars) <- c(0,0)
merc_do_Y200_cars <- crop(merc_do_Y200_cars, e)

diff_merc_cars200 <- diff(c(merc_do_Y200_cars, cars_do_resamp200))

d_do1.5_200 <- ggplot() + 
  geom_spatraster(data = diff_merc_cars200) + 
  ggtitle("Comparing Mercator to CSIRO") +
  labs(fill = "Difference (mmol/L)") + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()

#Merc vs. CalCOFI
vect_cal_do200 <- vect(calcofi_sta200)
df_cal_do200 <- as.data.frame(vect_cal_do200)
merc_cal_extract200 <- terra::extract(merc_do_Y200_cal, vect_cal_do200)
diff_merc_cal200 <- merc_cal_extract200$mean - df_cal_do200$mean_do200_mmol_L

calcofi_sta200$merc_diff200 <- diff_merc_cal200

d_do2_200 <- ggplot()+
  geom_sf(data = calcofi_sta200, aes(color = merc_diff200), size = 3)+
  scale_colour_gradientn(colours = matlab.like(30))+
  labs(color = "Difference (mmol/L)") + 
  ggtitle("Comparing Mercator to CalCOFI") +
  theme_minimal()

ggpubr::ggarrange(d_do1_200, d_do1.5_200, d_do2_200)
```

### Mercator temporal variation vs. CalCOFI data 200m

Here, the mercator dissolved oxygen rasters were subset the month-day-year that a CalCOFI bottle sample was collected. The plotted mercator values represent the closest grid cell to that CalCOFI station when a sample was collected. A daily temporal resolution was possible here.

```{r}
#| warning: false
#| message: false
#| echo: false
#| fig-align: center
#| fig-width: 12
#| fig-height: 4
#| label: temporal_variation200_DO

#Merc subset to CalCOFI domain in year-month temporal resolution 
merc_do_D200 <- merc_do200/1000
calcofi_coord200$Date <- as.Date(calcofi_coord200$Date, format = "%m/%d/%Y")
calcofi_coord200 <- calcofi_coord200 %>% drop_na(Date)

merc_vari200 <- NULL
for(i in 1:nrow(calcofi_coord200)){
  #subset raster according to calcofi sample date
  temp_date <- calcofi_coord200$Date[i]
  temp_rast <- merc_do_D200[[time(merc_do_D200) == temp_date]]
  
  #get coords of station 
  sta_loc <- calcofi_coord200$geometry[i]
  
  #extract mercator temp at the cell closest to that station
  sta_vect <- vect(sta_loc)
  merc_extract <- terra::extract(temp_rast, sta_vect)
  
  #save merc cell value, station value, and difference between in three cols
  temp_vari <- NULL
  temp_vari$Date <- temp_date
  temp_vari$geometry <- sta_loc
  temp_vari$merc_do_200m <- merc_extract[,2]
  temp_vari$calcofi_do_200m <- calcofi_coord200$do200_mmol_L[i]
  
  df_vari <- as.data.frame(temp_vari)
  
  merc_vari200 <- rbind(merc_vari200, df_vari)
}

merc_vari200_l <- merc_vari200 %>% pivot_longer(c(3:4), names_to = "data_source", values_to = "DO_mmol_L") %>%
  mutate(sta_geom = as.character(geometry))%>% st_as_sf()

#filter for coastal cluster of stations 
merc_vari200_coast <- st_crop(merc_vari200_l, xmin = -121, xmax = -117, ymax = 34, ymin = 31)
merc_vari200_coast <- merc_vari200_coast %>% 
  mutate(yr_mo = substr(Date, 1, 7)) %>%
  group_by(yr_mo, data_source) %>% 
  mutate(mean_DO = mean(DO_mmol_L, na.rm = TRUE))

ggplot(merc_vari200_coast, aes(Date, mean_DO, color = data_source)) + 
  geom_point(size = 2) + 
  geom_line() +
  scale_color_manual(values = c("steelblue4", "goldenrod"))+
  labs(subtitle = "Coastal stations") + 
  theme_minimal()

#filter for offshore cluster of stations 
merc_vari200_off <- st_crop(merc_vari200_l, xmin = -125, xmax = -121.5, ymax = 32.5, ymin = 28)
merc_vari200_off <- merc_vari200_off %>% 
  mutate(yr_mo = substr(Date, 1, 7)) %>%
  group_by(yr_mo, data_source) %>% 
  mutate(mean_DO = mean(DO_mmol_L, na.rm = TRUE))

ggplot(merc_vari200_off, aes(Date, mean_DO, color = data_source)) + 
  geom_point(size = 2) + 
  geom_line() +
  scale_color_manual(values = c("steelblue4", "goldenrod"))+
  labs(subtitle = "Offshore stations") + 
  theme_minimal()
```

## Temp (C) at 0m

### Mean across study period (2003-2015)

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: temp_0m
#| fig-align: center
#| fig-width: 10
#| fig-height: 8

#Mercator plot -- mean across all years
merc_temp_Y0_woa <- aggregate(merc_temp_Y0, fact = 4)
merc_temp_Y0_woa <- mean(merc_temp_Y0_woa, na.rm=TRUE)
merc_temp_Y0_cal <- mean(merc_temp_Y0, na.rm = TRUE)
merc_temp_Y0_cars <- aggregate(merc_temp_Y0, fact = 2)
merc_temp_Y0_cars <- mean(merc_temp_Y0_cars, na.rm = TRUE)

temp1_0 <- ggplot() + 
  geom_spatraster(data = merc_temp_Y0_cal) + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()+
    ggtitle("Mercator") +
  labs(fill = "Temp (C)")

#WOA plot -- mean across all years
temp2_0 <- ggplot() + 
  geom_spatraster(data = temp_woa0) + 
  scale_fill_gradientn(colours = matlab.like(30))+
   labs(fill = "Temp (C)")+
    ggtitle("WOA") +
  theme_minimal()

#CSIRO plot -- mean across all years
temp2.5_0 <- ggplot() + 
  geom_spatraster(data = cars_temp_resamp0) + 
  scale_fill_gradientn(colours = matlab.like(30))+
   labs(fill = "Temp (C)")+
    ggtitle("CSIRO") +
  theme_minimal()

#OISST plot -- mean across all years
clim_temp_Y0 <- mean(clim_temp_Y0, na.rm = TRUE)
temp3_0 <- ggplot() + 
  geom_spatraster(data = clim_temp_Y0) + 
  scale_fill_gradientn(colours = matlab.like(30))+
   labs(fill = "Temp (C)")+
    ggtitle("OISST") +
  theme_minimal()

#calcofi plot -- mean across all years for each station 
calcofi_coord0 <- st_as_sf(calcofi_samp_coords, coords = c("Lon_Deg", "Lat_Deg"), crs = st_crs(4326))

calcofi_sta0 <- calcofi_coord0 %>% 
  group_by(geometry) %>% 
  summarise(mean_temp0 = mean(T_degC, na.rm = TRUE))

temp4_0 <- ggplot()+
  geom_sf(data = calcofi_sta0, aes(color = mean_temp0), size = 3)+
  scale_colour_gradientn(colours = matlab.like(30))+
  theme_minimal()+
    ggtitle("CalCOFI") +
  labs(color = "Temp (C)")

ggpubr::ggarrange(temp1_0, temp2_0, temp2.5_0, temp3_0, temp4_0)
```

### Mercator difference plots

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: temp_diff_0m
#| fig-width: 10
#| fig-height: 8
#Merc vs. OISST
ext(clim_temp_Y0) <- c(-153, -104.25, 1.877594, 48.87759)
resamp_geom <- rast(ext(-153, -104.25, 1.877594, 48.87759), 
                    crs = "EPSG:4326",
                    res = 0.25)
OISST_resamp0 <- terra::resample(clim_temp_Y0, resamp_geom, method='bilinear')
merc_temp_Y0_cal <- crop(merc_temp_Y0_cal, e)
OISST_temp_Y0 <- crop(OISST_resamp0, e)

diff_merc_OISST0 <- diff(c(merc_temp_Y0_cal, OISST_temp_Y0))

d_temp1_0 <- ggplot() + 
  geom_spatraster(data = diff_merc_OISST0) + 
  ggtitle("Comparing Mercator to OISST") +
  labs(fill = "Difference (C)") + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()

#Merc vs. WOA
ext(temp_woa0) <- c(-153.75, -103.75, 0.377594, 51.37759)
resamp_geom_woa <- rast(ext(-153.75, -103.75, 0.377594, 51.37759), 
                    crs = "EPSG:4326",
                    res = 1)
woa_resamp0 <- terra::resample(temp_woa0, resamp_geom_woa, method='bilinear')

diff_merc_WOA0 <- diff(c(merc_temp_Y0_woa, woa_resamp0))

d_temp2_0 <- ggplot() + 
  geom_spatraster(data = diff_merc_WOA0) + 
  ggtitle("Comparing Mercator to WOA") +
  labs(fill = "Difference (C)") + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()

#Merc vs. CARS CSIRO
origin(merc_temp_Y0_cars) <- c(0,0)
merc_temp_Y0_cars <- crop(merc_temp_Y0_cars, e)

diff_merc_cars_temp0 <- diff(c(merc_temp_Y0_cars, cars_temp_resamp0))

d_temp2.5_0 <- ggplot() + 
  geom_spatraster(data = diff_merc_cars_temp0) + 
  ggtitle("Comparing Mercator to CSIRO") +
  labs(fill = "Difference (C)") + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()

#Merc vs. CalCOFI
vect_cal_temp0 <- vect(calcofi_sta0)
df_cal_temp0 <- as.data.frame(vect_cal_temp0)
merc_cal_extract0 <- terra::extract(merc_temp_Y0_cal, vect_cal_temp0)
diff_merc_cal0 <- merc_cal_extract0$mean - df_cal_temp0$mean_temp0

calcofi_sta0$merc_diff0 <- diff_merc_cal0

d_temp3_0 <- ggplot()+
  geom_sf(data = calcofi_sta0, aes(color = merc_diff0), size = 3)+
  scale_colour_gradientn(colours = matlab.like(30))+
  labs(color = "Difference (C)") + 
  ggtitle("Comparing Mercator to CalCOFI") +
  theme_minimal()

ggpubr::ggarrange(d_temp2_0, d_temp1_0, d_temp2.5_0, d_temp3_0)
```

### Mercator temporal variation vs. CalCOFI data 0m

Here, the mercator temperature rasters were subset the month-year that a CalCOFI bottle sample was collected. The plotted mercator values represent the closest grid cell to that CalCOFI station when a sample was collected. A monthly temporal resolution was possible here and not daily due to computational power of extracting the mercator temperature data directly from mercator rather than Copernicus, which is how the dissolved oxygen rasters were obtained.

```{r}
#| warning: false
#| message: false
#| echo: false
#| fig-align: center
#| fig-width: 12
#| fig-height: 6
#| label: temporal_variation0_Temp

calcofi_coord0_temp <- calcofi_coord0 %>% mutate(mo_ch = substr(Date, 1,2),
                                  yr_mo_day = paste0(Year,"-", mo_ch, "-", "01"),
                                  Date_merc = as.Date(yr_mo_day, format = "%Y-%m-%d"))

calcofi_coord0_temp <- calcofi_coord0_temp %>% drop_na(Date_merc)

merc_vari0_temp <- NULL
for(i in 1:nrow(calcofi_coord0_temp)){
  #subset raster according to calcofi sample date
  temp_date <- calcofi_coord0_temp$Date_merc[i]
  temp_rast <- merc_temp_all0[[time(merc_temp_all0) == temp_date]]

  #get coords of station
  sta_loc <- calcofi_coord0_temp$geometry[i]

  #extract mercator temp at the cell closest to that station
  sta_vect <- vect(sta_loc)
  merc_extract <- terra::extract(temp_rast, sta_vect)

  #save merc cell value, station value, and difference between in three cols
  temp_vari <- NULL
  temp_vari$Date <- temp_date
  temp_vari$geometry <- sta_loc
  temp_vari$merc_temp_0m <- merc_extract[,2]
  temp_vari$calcofi_temp_0m <- calcofi_coord0_temp$T_degC[i]

  df_vari <- as.data.frame(temp_vari)

  merc_vari0_temp <- rbind(merc_vari0_temp, df_vari)
}

merc_vari0_temp_l <- merc_vari0_temp %>% pivot_longer(c(3:4), names_to = "data_source", values_to = "Temp_C") %>%
  mutate(sta_geom = as.character(geometry)) %>% st_as_sf()


#filter for coastal cluster of stations 
merc_vari0_temp_coast <- st_crop(merc_vari0_temp_l, xmin = -121, xmax = -117, ymax = 34, ymin = 31)
merc_vari0_temp_coast <- merc_vari0_temp_coast %>% 
  mutate(yr_mo = substr(Date, 1, 7)) %>%
  group_by(yr_mo, data_source) %>% 
  mutate(mean_temp = mean(Temp_C, na.rm = TRUE))

ggplot(merc_vari0_temp_coast, aes(Date, mean_temp, color = data_source)) + 
  geom_point(size = 2) + 
  geom_line() +
  scale_color_manual(values = c("steelblue4", "goldenrod"))+
  labs(subtitle = "Coastal stations") + 
  theme_minimal()

#filter for offshore cluster of stations 
merc_vari0_temp_off <- st_crop(merc_vari0_temp_l, xmin = -125, xmax = -121.5, ymax = 32.5, ymin = 28)
merc_vari0_temp_off  <- merc_vari0_temp_off %>% 
  mutate(yr_mo = substr(Date, 1, 7)) %>%
  group_by(yr_mo, data_source) %>% 
  mutate(mean_temp = mean(Temp_C, na.rm = TRUE))

ggplot(merc_vari0_temp_off, aes(Date, mean_temp, color = data_source)) + 
  geom_point(size = 2) + 
  geom_line() +
  scale_color_manual(values = c("steelblue4", "goldenrod"))+
  labs(subtitle = "Offshore stations") + 
  theme_minimal()

```

## Temp (C) at 200m

### Mean across study period (2003-2015)

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: temp_200m
#| fig-width: 10
#| fig-height: 8
#Mercator plot -- mean across all years
merc_temp_Y200_woa <- aggregate(merc_temp_Y200, fact = 4)
merc_temp_Y200_woa <- mean(merc_temp_Y200_woa, na.rm=TRUE)
merc_temp_Y200_cal <- mean(merc_temp_Y200, na.rm = TRUE)
merc_temp_Y200_cars <- aggregate(merc_temp_Y200, fact = 2)
merc_temp_Y200_cars <- mean(merc_temp_Y200_cars, na.rm = TRUE)

temp1_200 <- ggplot() + 
  geom_spatraster(data = merc_temp_Y200_cal) + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()+
    ggtitle("Mercator") +
  labs(fill = "Temp (C)")

#WOA plot -- mean across all years
temp2_200 <- ggplot() + 
  geom_spatraster(data = temp_woa200) + 
  scale_fill_gradientn(colours = matlab.like(30))+
   labs(fill = "Temp (C)")+
    ggtitle("WOA") +
  theme_minimal()

#CSIRO plot -- mean across all years
temp2.5_200 <- ggplot() + 
  geom_spatraster(data = cars_temp_resamp200) + 
  scale_fill_gradientn(colours = matlab.like(30))+
   labs(fill = "Temp (C)")+
    ggtitle("CSIRO") +
  theme_minimal()

#calcofi plot -- mean across all years for each station 
calcofi_coord200 <- st_as_sf(calcofi_samp_coords200, coords = c("Lon_Deg", "Lat_Deg"), crs = st_crs(4326))

calcofi_sta200 <- calcofi_coord200 %>% 
  group_by(geometry) %>% 
  summarise(mean_temp200 = mean(T_degC, na.rm = TRUE))

temp3_200 <- ggplot()+
  geom_sf(data = calcofi_sta200, aes(color = mean_temp200), size = 3)+
  scale_colour_gradientn(colours = matlab.like(30))+
  theme_minimal()+
    ggtitle("CalCOFI") +
  labs(color = "Temp (C)")

ggpubr::ggarrange(temp1_200, temp2_200, temp2.5_200, temp3_200)
```

### Mercator difference plots

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: temp_diff_200m
#| fig-width: 10
#| fig-height: 8
#Merc vs. WOA
woa_resamp200 <- terra::resample(temp_woa200, resamp_geom_woa, method='bilinear')
diff_merc_WOA200 <- diff(c(merc_temp_Y200_woa, woa_resamp200))

d_temp1_200 <- ggplot() + 
  geom_spatraster(data = diff_merc_WOA200) + 
  ggtitle("Comparing Mercator to WOA") +
  labs(fill = "Difference (C)") + 
  scale_fill_gradientn(colours = matlab.like(30))+ 
  theme_minimal()

#Merc vs. CARS CSIRO
origin(merc_temp_Y200_cars) <- c(0,0)
merc_temp_Y200_cars <- crop(merc_temp_Y200_cars, e)

diff_merc_cars_temp200 <- diff(c(merc_temp_Y200_cars, cars_temp_resamp200))

d_temp1.5_200 <- ggplot() + 
  geom_spatraster(data = diff_merc_cars_temp200) + 
  ggtitle("Comparing Mercator to CSIRO") +
  labs(fill = "Difference (C)") + 
  scale_fill_gradientn(colours = matlab.like(30))+
  theme_minimal()

#Merc vs. CalCOFI
vect_cal_temp200 <- vect(calcofi_sta200)
df_cal_temp200 <- as.data.frame(vect_cal_temp200)
merc_cal_extract200 <- terra::extract(merc_temp_Y200_cal, vect_cal_temp200)
diff_merc_cal200 <- merc_cal_extract200$mean - df_cal_temp200$mean_temp200

calcofi_sta200$merc_diff200 <- diff_merc_cal200

d_temp2_200 <- ggplot()+
  geom_sf(data = calcofi_sta200, aes(color = merc_diff200), size = 3)+
  scale_colour_gradientn(colours = matlab.like(30))+
  labs(color = "Difference (C)") + 
  ggtitle("Comparing Mercator to CalCOFI") +
  theme_minimal()

ggpubr::ggarrange(d_temp1_200, d_temp1.5_200, d_temp2_200)
```

### Mercator temporal variation vs. CalCOFI data 200m

Here, the mercator temperature rasters were subset the month-year that a CalCOFI bottle sample was collected. The plotted mercator values represent the closest grid cell to that CalCOFI station when a sample was collected. A monthly temporal resolution was possible here and not daily due to computational power of extracting the mercator temperature data directly from mercator rather than Copernicus, which is how the dissolved oxygen rasters were obtained.

```{r}
#| warning: false
#| message: false
#| echo: false
#| fig-align: center
#| fig-width: 12
#| fig-height: 6
#| label: temporal_variation200_Temp

calcofi_coord200_temp <- calcofi_coord200 %>% mutate(mo_ch = substr(Date, 1,2),
                                  yr_mo_day = paste0(Year, "-", mo_ch, "-", "01"),
                                  Date_merc = as.Date(yr_mo_day, format = "%Y-%m-%d"))
calcofi_coord200_temp <- calcofi_coord200_temp %>% drop_na(Date_merc)
merc_vari200_temp <- NULL
for(i in 1:nrow(calcofi_coord200_temp)){
  #subset raster according to calcofi sample date
  temp_date <- calcofi_coord200_temp$Date_merc[i]
  temp_rast <- merc_temp_all200[[time(merc_temp_all200) == temp_date]]

  #get coords of station
  sta_loc <- calcofi_coord200_temp$geometry[i]

  #extract mercator temp at the cell closest to that station
  sta_vect <- vect(sta_loc)
  merc_extract <- terra::extract(temp_rast, sta_vect)

  #save merc cell value, station value, and difference between in three cols
  temp_vari <- NULL
  temp_vari$Date <- temp_date
  temp_vari$geometry <- sta_loc
  temp_vari$merc_temp_200m <- merc_extract[,2]
  temp_vari$calcofi_temp_200m <- calcofi_coord200_temp$T_degC[i]

  df_vari <- as.data.frame(temp_vari)

  merc_vari200_temp <- rbind(merc_vari200_temp, df_vari)
}

merc_vari200_temp_l <- merc_vari200_temp %>% pivot_longer(c(3:4), names_to = "data_source", values_to = "Temp_C") %>%
  mutate(sta_geom = as.character(geometry)) %>% st_as_sf()

#filter for coastal cluster of stations 
merc_vari200_temp_coast <- st_crop(merc_vari200_temp_l, xmin = -121, xmax = -117, ymax = 34, ymin = 31)
merc_vari200_temp_coast <- merc_vari200_temp_coast %>% 
  mutate(yr_mo = substr(Date, 1, 7)) %>%
  group_by(yr_mo, data_source) %>% 
  mutate(mean_temp = mean(Temp_C, na.rm = TRUE))

ggplot(merc_vari200_temp_coast, aes(Date, mean_temp, color = data_source)) + 
  geom_point(size = 2) + 
  geom_line() +
  scale_color_manual(values = c("steelblue4", "goldenrod"))+
  labs(subtitle = "Coastal stations") + 
  theme_minimal()

#filter for offshore cluster of stations 
merc_vari200_temp_off <- st_crop(merc_vari200_temp_l, xmin = -125, xmax = -121.5, ymax = 32.5, ymin = 28)
merc_vari200_temp_off  <- merc_vari200_temp_off %>% 
  mutate(yr_mo = substr(Date, 1, 7)) %>%
  group_by(yr_mo, data_source) %>% 
  mutate(mean_temp = mean(Temp_C, na.rm = TRUE))

ggplot(merc_vari200_temp_off, aes(Date, mean_temp, color = data_source)) + 
  geom_point(size = 2) + 
  geom_line() +
  scale_color_manual(values = c("steelblue4", "goldenrod"))+
  labs(subtitle = "Offshore stations") + 
  theme_minimal()
```

